---
title: "Emotive speech nmf analyses""
author: "Helio"
date: "2024-08-23"
output: html_document
---

dependencies: preprocessing.Rmd
preferences
```{r}
library(tidyverse)
library(data.table)
library(NMF)

```

for the NMF we will start with k = 3 given our findings in the posed (we have a good prior for 3). 
But then we will also validate the k


```{r}
# use only the the emotive speech condition called "spoken" not spojen 2
unique(df_OF_output_AUsW_unblind_spoken1_2_binned$filename)

df_OF_output_AUsW_unblind_spoken1_binned<- subset(df_OF_output_AUsW_unblind_spoken1_2_binned,
                                           posed.spoken == "spoken")


```

Fit NMF


```{r}


# range(df_OF_output_AUsW_unblind_spoken1_binned$AU04_r_Brow_lowerer)

colnames(df_OF_output_AUsW_unblind_spoken1_binned)

df_OF_output_AUsW_unblind_spoken1_binned$test.day
select_au_and_id_cols<- colnames(df_OF_output_AUsW_unblind_spoken1_binned[,c(1:4,10,17:33)])

# select_au_and_id_cols
#  [1] "filename"                    "bin_frame"                  
#  [3] "subject"                     "test.day"                   
#  [5] "expression"                  "AU01_r_Inner_brow_raiser"   
#  [7] "AU02_r_Outer_brow_raiser"    "AU04_r_Brow_lowerer"        
#  [9] "AU05_r_Upper_lid_raiser"     "AU06_r_Cheek_raiser"        
# [11] "AU07_r_Lid_tightener"        "AU09_r_Nose_wrinkler"       
# [13] "AU10_r_Upper_lip_raiser"     "AU12_r_Lip_corner_puller"   
# [15] "AU14_r_Dimpler"              "AU15_r_Lip_corner_depressor"
# [17] "AU17_r_Chin_raiser"          "AU20_r_Lip_stretcher"       
# [19] "AU23_r_Lip_tightener"        "AU25_r_Lips_part"           
# [21] "AU26_r_Jaw_drop"             "AU45_r_Blink" 


# we need to store with some identifiers
saveRDS(df_OF_output_AUsW_unblind_spoken1_binned[,select_au_and_id_cols],
        file.path(paths$export_data, 
                  "dta_spk_4nmf.rds"))

# dta_spk_4nmf.rds data for emotive speech NMF on whole dataset

res_k3_spoken1 <- NMF::nmf(df_OF_output_AUsW_unblind_spoken1_binned[,17:33], r = 3, 
                  nrun = 200, #number of runs to try and update for
                  seed=123456, #specific seed for reproducibility
                 .options = list( 'v')) #verbose


saveRDS(res_k3_spoken1,
        file.path(paths$export_models, 
                  "mod_rslt_nmfspk.rds"))

# mod_rslt_nmfspk.rds NMF model results on the whole emotive speech data.


range(df_OF_output_AUsW_unblind_spoken1_binned$bin_frame)
df_OF_output_AUsW_unblind_spoken1_binned$AU25_r_Lips_part

require(NMF)

# train and test from the nmf stage

# createDataPartition()
```


 NMF SPOKEN: permutation validation (find k)


```{r}
res_spoken1_find_k <- nmf(df_OF_output_AUsW_unblind_spoken1_binned[,17:33]+.0001, 2:6, 
                      seed=123456,
                      nrun = 100,
                       .options = 'v')


# memory exhaust - takes a lot of time
```


colnames(dta_spk_4nmf)
```{r}
set.seed(123456)
# Define the size of the subset (e.g., 50% of the original data)
subset_size <- 0.5

# Create a vector of unique timeseries IDs
spk_unique_timeseries_ids <- unique(df_OF_output_AUsW_unblind_spoken1_binned$filename)

# Helper function to perform stratified sampling


# Perform stratified sampling to select unique timeseries IDs based on the 'expression' column
unique(df_OF_output_AUsW_unblind_spoken1_binned$filename)
spk_stratified_ids <- df_OF_output_AUsW_unblind_spoken1_binned %>%
  group_by(expression) %>%
  summarise(filename = list(unique(filename))) %>%
  unnest(filename) %>%
  group_by(expression) %>%
  mutate(sample_id = row_number()) %>%
  ungroup() %>%
  filter(sample_id <= subset_size * n() / length(unique(expression))) %>%
  pull(filename)

# save.image("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/halofacestudyOCT2023.RData")
# Split the data based on the selected timeseries IDs
dta_subs_a_spk <- df_OF_output_AUsW_unblind_spoken1_binned %>%
  filter(filename %in% spk_stratified_ids)

table(dta_subs_a_spk$expression)

dta_subs_b_spk <- df_OF_output_AUsW_unblind_spoken1_binned %>%
  filter(!filename %in% spk_stratified_ids)

table(dta_subs_b_spk$expression)

colnames(dta_subs_a_spk)
dta_subs_a_spk
library(NMF)
dta_subs_a_spk

res_find_k_spk_subs <- NMF::nmf(as.matrix(dta_subs_a_spk[,17:33]+.001), 2:6, 
                      seed=123456,
                      nrun = 10,
                       .options = 'v')



```

block shuffling

```{r}
dta_spk_subs_a_rnd<- fn_temporal_block_permute(dta_subs_a_spk[,17:33], 
                                                          dta_subs_a_spk$filename,
                                                     time_block_size = 5) # using 5 as spech is more dynamic

# get attributes
dta_spk_subs_a_rnd$filename <- dta_subs_a_spk$filename
dta_spk_subs_a_rnd$expression <- dta_subs_a_spk$expression
dta_spk_subs_a_rnd$timestamp <- dta_subs_a_spk$timestamp


dta_subs_a_spk%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)


dta_spk_subs_a_rnd%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)



# estimate quality measures from the shuffled data (use default NMF algorithm)
colnames(dta_spk_subs_a_rnd)

estim_spk_random_1 <- NMF::nmf(dta_spk_subs_a_rnd[,1:17]+.001, 
                               2:6, nrun=10,
                               seed=123456,
                                .options = 'v')


plot(res_find_k_spk_subs)
plot(estim_spk_random_1)


plot(res_find_k_spk_subs, estim_spk_random_1)+
  geom_vline(xintercept = 3)

# plot fir comparisons
plots_SM$spk_fit_vs_rand <- plot(res_find_k_spk_subs,estim_spk_random_1)+
  geom_vline(xintercept = 3)


```



<!-- simplify -->


```{r}

NMF::coefmap(res_k3_spoken1)

spoken_NMF<- df_OF_output_AUsW_unblind_spoken1_binned[,c(1,2:3,9:10,34:36)]

# spoken_NMF_backup<-spoken_NMF

# length(spoken_NMF$k3_comp2)== length(res_k3_spoken1@fit@W[,1]) #TRUE
# just reordering based on how we describe in the paper and to match the order of the posed (expression only NMF)
spoken_NMF$k3_comp1<- res_k3_spoken1@fit@W[,2]
spoken_NMF$k3_comp2<- res_k3_spoken1@fit@W[,1]
spoken_NMF$k3_comp3<- res_k3_spoken1@fit@W[,3]

colnames(spoken_NMF)



```





extra variability of AU timecpourses

# sd
df_with_clusters6$posed.spoken<- "posed"
bind_rows(
  df_OF_output_AUsW_unblind_spoken1_binned [,c("filename","bin_frame", "subject" ,"expression","posed.spoken","drug.placebo", colnames(df_OF_output_AUsW_unblind_spoken1_binned[,c(17:33)]))] ,
  df_with_clusters6[,c("filename","bin_frame", "subject" ,"expression","posed.spoken","drug.placebo", colnames(df_OF_output_AUsW_unblind_spoken1_binned[,c(17:33)]))])%>%
  filter(!subject %in% tmp_subj_to_remove)%>%
    subset(expression!= "neutral")%>%
  # group_by(filename, subject,bin_frame, expression, posed.spoken)%>%


  gather("AU", "Coef", -bin_frame,-filename, -expression,-subject,-posed.spoken,-drug.placebo) %>%
      group_by(subject,posed.spoken,AU)%>%
  mutate(Coef = scale(as.numeric(Coef, na.rm = T))) %>%
  group_by(bin_frame, expression, posed.spoken,AU)%>%
  subset(!is.na(Coef))%>%
    summarise_at(c("Coef"), sd, na.rm = T) %>%

  ggplot(aes(bin_frame,AU,fill= Coef))+
  geom_tile()+
  facet_grid(expression~posed.spoken)+
  scale_fill_viridis_c(option = "magma")+
  ylab("Variability - SD (z)")

