---
title: "expression only nmf analyses"
author: "Helio"
date: "2024-08-23"
output: html_document
---

This script fits the spatiotemporal NMF for expression only whole data data in a few steps

1. initialisation = random seed
2. find k
3. fit final
3. Compare NMF to randomined NMF

dependencies: needs df_OF_output_AUsW_unblind_posed_binned from preprocessing.Rmd
preferences
```{r}
# install these is not already installed
library(tidyverse)
library(data.table)
library(NMF)
require(ggforce)
library(magick)
require(ggtext)
library(caret)
require(ggside)
require(ggExtra)
```


# Initialise and find k
```{r}
colnames(df_OF_output_AUsW_unblind_posed_binned)

# single run and single numeric seed for reproducibility
res_find_k <- nmf(df_OF_output_AUsW_unblind_posed_binned[,16:32], 2:6, 
                      seed=123456,
                      nrun = 100,
                       .options = 'v')




NMF::plot(res_find_k)
  geom_vline(xintercept = 3)

test$data
  
NMF::consensusmap(res_find_k)

```

2. NMF POSED: permutation validation
We select values of k where the magnitude of the cophenetic correlation coefficient begins to fall (see below).
so now we can fit the final model

between 3 and 4

Check for overfiting by reshufling the ata and refitting the NMF on simulated data
SHUFFLING to avoid overfiting to noise

```{r}

# Define the size of the subset (e.g., 50% of the original data)
subset_size <- 0.5

# Create a vector of unique timeseries IDs
unique_timeseries_ids <- unique(df_OF_output_AUsW_unblind_posed_binned$filename)

# Helper function to perform stratified sampling
stratified_sampling <- function(data, group_col, size) {
  data %>%
    group_by(!!sym(group_col)) %>%
    sample_frac(size = size / n(), replace = FALSE) %>%
    ungroup()
}

# Perform stratified sampling to select unique timeseries IDs based on the 'expression' column
stratified_ids <- df_OF_output_AUsW_unblind_posed_binned %>%
  group_by(expression) %>%
  summarise(filename = list(unique(filename))) %>%
  unnest(filename) %>%
  group_by(expression) %>%
  mutate(sample_id = row_number()) %>%
  ungroup() %>%
  filter(sample_id <= subset_size * n() / length(unique(expression))) %>%
  pull(filename)

# Split the data based on the selected timeseries IDs
dta_subs_a <- df_OF_output_AUsW_unblind_posed_binned %>%
  filter(filename %in% stratified_ids)

dta_subs_a
```




```{r}
dta_subs_b <- df_OF_output_AUsW_unblind_posed_binned %>%
  filter(!filename %in% stratified_ids)

colnames(dta_subs_a)

res_find_k_1 <- NMF::nmf(dta_subs_a[,16:32], 2:6, 
                      seed=123456,
                      nrun = 10,
                       .options = 'v')



```



```{r}


# Function for temporal block permutation with fixed block size - fixed is better as the shuflign wont overfit the data
fn_temporal_block_permute <- function(X, groups, time_block_size) {
  X_permuted <- X
  unique_groups <- unique(groups)
  
  for (group in unique_groups) {
    indices <- which(groups == group)
    n_samples <- length(indices)
    
    if (n_samples > 1) {
      # Perform block permutation for each column independently
      n_blocks <- floor(n_samples / time_block_size)
      if (n_blocks > 1) {
        for (col in 1:ncol(X)) {
          block_indices <- split(1:n_samples, rep(1:n_blocks, each = time_block_size, length.out = n_samples))
          permuted_order <- unlist(sample(block_indices))
          X_permuted[indices, col] <- X[indices[permuted_order], col]
        }
      } else {
        # If only one block, permute all time points for each column
        for (col in 1:ncol(X)) {
          X_permuted[indices, col] <- X[indices[sample(n_samples)], col]
        }
      }
    }
  }
  
  return(X_permuted)
}

# we can do a few time blocks
fn_temporal_block_permute()


 dta_posed_subs_a_rnd_2_tbs_5<- fn_temporal_block_permute(dta_subs_a[,16:32], 
                                                          dta_subs_a$filename,
                                                     time_block_size = 5)
 
dta_posed_subs_a_rnd_2_tbs_10<- fn_temporal_block_permute(dta_subs_a[,16:32], 
                                                          dta_subs_a$filename,
                                                     time_block_size = 10)
  
  
  dta_posed_subs_a_rnd_2_tbs_15<- fn_temporal_block_permute(dta_subs_a[,16:32], 
                                                          dta_subs_a$filename,
                                                     time_block_size = 15)


dta_posed_subs_a_rnd_2_tbs_5$filename <- dta_subs_a$filename
dta_posed_subs_a_rnd_2_tbs_5$expression <- dta_subs_a$expression
dta_posed_subs_a_rnd_2_tbs_5$timestamp <- dta_subs_a$timestamp


dta_posed_subs_a_rnd_2_tbs_10$filename <- dta_subs_a$filename
dta_posed_subs_a_rnd_2_tbs_10$expression <- dta_subs_a$expression
dta_posed_subs_a_rnd_2_tbs_10$timestamp <- dta_subs_a$timestamp

dta_posed_subs_a_rnd_2_tbs_15$filename <- dta_subs_a$filename
dta_posed_subs_a_rnd_2_tbs_15$expression <- dta_subs_a$expression
dta_posed_subs_a_rnd_2_tbs_15$timestamp <- dta_subs_a$timestamp



dta_subs_a%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)

dta_posed_subs_a_rnd_2_tbs_5%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)

dta_posed_subs_a_rnd_2_tbs_10%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)

dta_posed_subs_a_rnd_2_tbs_15%>%
  ggplot(aes(timestamp,AU06_r_Cheek_raiser, group = filename))+
  geom_line(colour = "blue", alpha = .5)+
  geom_line(aes(y = AU01_r_Inner_brow_raiser), colour = "red", alpha = .1)+
  facet_grid(~expression)

```

estimate NMF on randomised data

```{r}
# estimate quality measures from the shuffled data (use default NMF algorithm)
estim_posed_random_2_tbs_10 <- NMF::nmf(dta_posed_subs_a_rnd_2_tbs_10[,1:17], 
                                 2:6, nrun=10, seed=123456)

estim_posed_random_2_tbs_5 <- NMF::nmf(dta_posed_subs_a_rnd_2_tbs_5[,1:17], 
                                 2:6, nrun=10, seed=123456)


estim_posed_random_2_tbs_15 <- NMF::nmf(dta_posed_subs_a_rnd_2_tbs_15[,1:17], 
                                 2:6, nrun=10, seed=123456)


# save manually

```


NMF POSED: FINAL NMF

```{r}

colnames(df_OF_output_AUsW_unblind_posed_binned)
df_OF_output_AUsW_unblind_posed_binned$AU01_r_Inner_brow_raiser


# for OSF
col_names_aus_4nmf<- colnames(df_OF_output_AUsW_unblind_posed_binned[,c(1:4,10,16:32)])



# see localpaths.Rmd in private
saveRDS(df_OF_output_AUsW_unblind_posed_binned[,c(1:4,10,16:32)],
         file.path(paths$export_data, 
                  "dta_expr_only_4nmf.rds"))

# dta_expr_only_4nmf.rds contains expression only AU data for NMF fitting
# colnames(df_OF_output_AUsW_unblind_posed_binned[,c(1:4,10,16:32)])
# [1] "filename"                    "bin_frame"                  
#  [3] "subject"                     "expression"                 
#  [5] "AU01_r_Inner_brow_raiser"    "AU02_r_Outer_brow_raiser"   
#  [7] "AU04_r_Brow_lowerer"         "AU05_r_Upper_lid_raiser"    
#  [9] "AU06_r_Cheek_raiser"         "AU07_r_Lid_tightener"       
# [11] "AU09_r_Nose_wrinkler"        "AU10_r_Upper_lip_raiser"    
# [13] "AU12_r_Lip_corner_puller"    "AU14_r_Dimpler"             
# [15] "AU15_r_Lip_corner_depressor" "AU17_r_Chin_raiser"         
# [17] "AU20_r_Lip_stretcher"        "AU23_r_Lip_tightener"       
# [19] "AU25_r_Lips_part"            "AU26_r_Jaw_drop"            
# [21] "AU45_r_Blink" 

res_k3 <- NMF::nmf(df_OF_output_AUsW_unblind_posed_binned[,16:32], r = 3, 
                  nrun = 200, 
                  seed=123456,
                 .options = list( 'v')) #

res_k3@misc
# coefmap(res_k3)
# basismap(res_k3)

```


store NMF components on dataset
```{r}


df_OF_output_AUsW_unblind_posed_binned$NMFtable<- as.data.frame(res_k3@fit@W)

df_OF_output_AUsW_unblind_posed_binned$comp1 = df_OF_output_AUsW_unblind_posed_binned$NMFtable$V1
table(is.na(df_OF_output_AUsW_unblind_posed_binned$comp1))
df_OF_output_AUsW_unblind_posed_binned$comp2 = df_OF_output_AUsW_unblind_posed_binned$NMFtable$V2
df_OF_output_AUsW_unblind_posed_binned$comp3 = df_OF_output_AUsW_unblind_posed_binned$NMFtable$V3

```




```{r}

df_OF_output_AUsW_unblind_posed_binned
saveRDS(res_k3, "mod_rslt_nmfposed.Rds")
# mod_rslt_nmfposed.Rds object that contains expression only NMF results on the whole data

# HaloStudy/Data/ExportedData/Processeddata"
# My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Processeddata
saveRDS(df_OF_output_AUsW_unblind_posed_binned, "df_OF_output_AUsW_unblind_posed_binned.Rds")


```

final outputs

res_k3
df_OF_output_AUsW_unblind_posed_binned
res_find_k






