---
title: "Facial expression substate analyses"
author: "Helio"
output: html_document
---



```{r}

library(dplyr)
library(tsclust)
# install.packages("sjstats")
# install.packages("sjPlot")
# install.packages("psycho")
# install.packages("remotes")
# install.packages("easystats")


```


```{r}

# Define a custom function to perform max-norm normalization
maxnorm <- function(x) {
  x / max(abs(x))
}

# Define a custom function to perform clustering with tsclust
perform_tsclust_clustering <- function(data) {
  tsclust_result <- tsclust(data[, c("sp_avg", "comp_delta_avg", "compavg")], k = 3)
  return(tsclust_result@cluster)
}

# Apply max-norm normalization and perform tsclust clustering
unique(df_OF_output_AUsW_unblind_posed_binned3_nona_clust$sp_avg)
range(df_OF_output_AUsW_unblind_posed_binned3_nona_clust$sp_avg)
df_with_clusters6 <- df_OF_output_AUsW_unblind_posed_binned3_nona_clust %>%
  group_by(filename) %>%
    # group_by(subject) %>%
  mutate(sp_avg = maxnorm(sp_avg),
         comp_delta_avg = maxnorm(comp_delta_avg),
         compavg = maxnorm(compavg)) %>%
  ungroup() %>%
  mutate(cluster_across = perform_tsclust_clustering(cur_data()))

# quick check

df_with_clusters6%>%
  # subset(filename == "./cut_posed_sad_day1_p7.csv")%>%

  ggplot(aes(as.numeric(bin_frame), as.factor(subject), fill = as.factor(cluster_across)))+
  geom_tile()+

   scale_fill_viridis_d(option = "magma")+

  facet_grid(~expression)+
  theme_classic()

```


# illustration


```{r}
df_with_clusters6%>%
  subset(idno == 30)%>%

  ggplot(aes(as.numeric(bin_frame), sp_avg))+

  geom_rect(aes(xmin = bin_frame, xmax = dplyr::lag(bin_frame), ymin = -Inf, ymax = Inf, fill = factor(cluster_across)), 
            alpha = .6)  +
    geom_line(size = 1)+
  geom_line(aes(y = comp_delta_avg),  size = 1, linetype = 2)+
  geom_line(aes(y = comp1))+
  geom_line(aes(y = comp2))+
  geom_line(aes(y = comp3))+
  # geom_rect(aes(xmin=kmean_order-0.5, xmax=kmean_order+0.5))+ 
  # geom_smooth(se = F)+
  # facet_grid(~drug.placebo)+
  theme_classic()



```

# name clusters

```{r}
df_with_clusters6$cluster_label <- if_else(df_with_clusters6$cluster_across == 2,
                                             "relaxed",
                                             if_else(df_with_clusters6$cluster_across == 1,
                                             "transition",
                                             if_else(df_with_clusters6$cluster_across == 3,
                                             "sustain",NA)))




colnames(df_with_clusters6)
df_with_clusters6%>%
  mutate(speed = sp_avg,
         displacement = comp_delta_avg,
         nmf_k1 = comp1,
         nmf_k2 = comp2,
         nmf_k3 = comp3)%>%
  subset(idno == 10)%>%
  select(c(bin_frame,speed,displacement,nmf_k1,nmf_k2,nmf_k3,cluster_label))%>%
  # gather(variable,value,-bin_frame,-cluster_label,-filename)%>%

  # test
  gather(variable, coef,-bin_frame,-filename,-cluster_label) %>%
  mutate(var_type = if_else(grepl("speed", variable), "2. speed",  
                                
                                        if_else(grepl("displacement", variable), "3. displacement",  
 if_else(grepl("nmf", variable ), "1. nmf",NA)))) %>%

group_by(var_type)%>%
  mutate(coef_z = scale(coef))%>%
  # group_by(variable, bin_frame,var_type, cluster_label)%>%
  # summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ggplot(aes(bin_frame, coef_z, group = variable))+
  
    geom_rect(aes(xmin = bin_frame, xmax = dplyr::lag(bin_frame), ymin = -Inf, ymax = Inf, fill = cluster_label),
            alpha = .3)  +
  geom_line(aes(group = variable, linetype = var_type), size = 1)+
   facet_grid(var_type~.)+
  # facet_grid(variable_type~.)+
  theme_classic()

```



SUBSTATE POSED
spoken_NMF

```{r}
colnames(spoken_NMF)

unique(spoken_NMF$posed_spoken)

require("zoo")
spoken_NMF<- spoken_NMF %>%
  group_by(filename)%>%
  ungroup()%>%
  # subset(filename == "./cut_posed_angry_day1_p1.csv")%>%
  # select(c(2,38,42))%>%
  mutate(k3_comp1_speed = (abs(lag(k3_comp1) - k3_comp1))/abs((lead(bin_frame)-bin_frame)),
         k3_comp2_speed = (abs(lag(k3_comp2) - k3_comp2))/abs((lead(bin_frame)-bin_frame)),
         k3_comp3_speed = (abs(lag(k3_comp3) - k3_comp3))/abs((lead(bin_frame)-bin_frame)),
        
         k3_comp1_diff = as.numeric(diff(zoo::zoo(k3_comp1), na.pad = TRUE)),
         k3_comp2_diff =as.numeric(diff(zoo::zoo(k3_comp2), na.pad = TRUE)),
         k3_comp3_diff =as.numeric(diff(zoo::zoo(k3_comp3), na.pad = TRUE)),
         
         # moving averages
         k3_comp1_diff_ma = as.numeric(zoo::rollmean(k3_comp1_diff, k = 5, fill = NA)),
         k3_comp2_diff_ma = as.numeric(zoo::rollmean(k3_comp2_diff, k = 5, fill = NA)),
         k3_comp3_diff_ma = as.numeric(zoo::rollmean(k3_comp3_diff, k = 5, fill = NA)),
           
          k3_comp1_speed_ma = as.numeric(zoo::rollmean(k3_comp1_speed, k = 5, fill = NA)),
        k3_comp2_speed_ma = as.numeric(zoo::rollmean(k3_comp2_speed, k = 5, fill = NA)),
         k3_comp3_speed_ma = as.numeric(zoo::rollmean(k3_comp3_speed, k = 5, fill = NA)))%>%
  
  
  mutate(sp_avg = (k3_comp1_speed_ma+k3_comp2_speed_ma+k3_comp3_speed_ma)/3)%>%
   mutate(comp_delta_avg = (k3_comp1_diff_ma +k3_comp2_diff_ma+k3_comp3_diff_ma)/3)%>%
  
  
  mutate(peak_speed =  sum(abs(sp_avg), na.rm = TRUE),
           max_displacement = sum(abs(comp_delta_avg), na.rm = TRUE),
          compavg = sum(abs(k3_comp1+k3_comp2+k3_comp3), na.rm = TRUE)
         )


# plot to see if we can viusally notice subates based on speed profiles
unique(spoken_NMF$filename)
spoken_NMF$filename




```
  
# cluster substates here spoken
  
```{r}
# Load the required packages

# Define a custom function to perform clustering with tsclust
perform_tsclust_clustering <- function(data) {
  tsclust_result <- tsclust(data[, c("sp_avg", "comp_delta_avg", "compavg")], k = 3)
  return(tsclust_result@cluster)
}

# Apply max-norm normalization and perform tsclust clustering

?zoo::na.approx()
library(zoo)

spoken_NMF_int <- spoken_NMF %>%
  group_by(filename) %>%
  mutate(sp_avg = na.approx(sp_avg, na.rm = FALSE),
         comp_delta_avg = na.approx(comp_delta_avg, na.rm = FALSE),
 compavg = na.approx(compavg, na.rm = FALSE))%>%
  ungroup()


table(is.na(spoken_NMF_int$sp_avg))

table(is.na(spoken_NMF_int$comp_delta_avg))
table(is.na(spoken_NMF_int$compavg))

spoken_NMF_int$sp_avg<- if_else(is.na(spoken_NMF_int$sp_avg), 0,spoken_NMF_int$sp_avg)
spoken_NMF_int$comp_delta_avg<- if_else(is.na(spoken_NMF_int$comp_delta_avg), 0,spoken_NMF_int$comp_delta_avg)
spoken_NMF_int$compavg<- if_else(is.na(spoken_NMF_int$compavg), 0,spoken_NMF_int$compavg)

# rm(df_with_clusters5hier, df_with_clusters6_lm_nonl)
# rm(df_with_clusters3_agg)
# rm(dcast_time)
# rm(df_OF_output_AU_andLand)
# rm(df_OF_output_AUstest)
# 
# rm(df_OF_output_AUsW_unblind)
# rm(df_OF_output_land_posed_binned_select)
# rm(df_sp)

unique(spoken_NMF_int$drug.placebo)

# remove temporarfile.show
# rm(tmp_tbl, test, test_bind_agg_spoken_posed, test_bind_agg_spoken_posed_noneutr,
#    test_del, test_gradient,testspokenmtsnafilled, test_happy_dcast_summ, test_happy_dcast_summ_paired_test,
#    test4_df)
# rm(spoken_NMF_with_clusters6)

spoken_NMF_with_clusters6 


spoken_NMF_int<- spoken_NMF_int %>%
  group_by(filename) %>%
  # subset(drug.placebo == "placebo")%>%
  mutate(sp_avg = maxnorm(sp_avg),
         comp_delta_avg = maxnorm(comp_delta_avg),
         compavg = maxnorm(compavg))

# try to separate placebo and drug to avoid exhaust memory
spoken_NMF_with_clusters6_plac<- spoken_NMF_int%>%
  subset(drug.placebo == "placebo")%>%
  ungroup() %>%
  mutate(cluster_across = perform_tsclust_clustering(cur_data()))

# drug
spoken_NMF_int$drug.placebo

spoken_NMF_with_clusters6_drug<- spoken_NMF_int%>%
  subset(drug.placebo == "drug")%>%
  ungroup() %>%
  mutate(cluster_across = perform_tsclust_clustering(cur_data()))

# quick check

spoken_NMF_with_clusters6_drug%>%
  # subset(filename == "./cut_posed_sad_day1_p7.csv")%>%

  ggplot(aes(as.numeric(bin_frame), as.factor(subject), fill = as.factor(cluster_across)))+
  geom_tile()+

   scale_fill_viridis_d(option = "magma")+

  facet_grid(~expression)+
  theme_classic()

spoken_NMF_with_clusters6_plac%>%
  # subset(filename == "./cut_posed_sad_day1_p7.csv")%>%

  ggplot(aes(as.numeric(bin_frame), as.factor(subject), fill = as.factor(cluster_across)))+
  geom_tile()+

   scale_fill_viridis_d(option = "magma")+

  facet_grid(~expression)+
  theme_classic()
# 2 in drug is 3 in placebo
# 1 and 1 are the same?
# 3 is 2
?recode
spoken_NMF_with_clusters6_drug<- spoken_NMF_with_clusters6_drug%>%
  mutate(cluster_across_recode = recode(cluster_across, `2` = 3, `1` = 1, `3` =2))

# now copare again

spoken_NMF_with_clusters6_drug%>%
  # subset(filename == "./cut_posed_sad_day1_p7.csv")%>%

  ggplot(aes(as.numeric(bin_frame), as.factor(subject), fill = as.factor(cluster_across_recode)))+
  geom_tile()+

   scale_fill_viridis_d(option = "magma")+

  facet_grid(~expression)+
  theme_classic()

spoken_NMF_with_clusters6_plac%>%
  # subset(filename == "./cut_posed_sad_day1_p7.csv")%>%

  ggplot(aes(as.numeric(bin_frame), as.factor(subject), fill = as.factor(cluster_across)))+
  geom_tile()+

   scale_fill_viridis_d(option = "magma")+

  facet_grid(~expression)+
  theme_classic()

spoken_NMF_with_clusters6_drug$cluster_across<- spoken_NMF_with_clusters6_drug$cluster_across_recode
spoken_NMF_with_clusters6_drug$cluster_across_recode<- NULL

# now merge drug and placebo
spoken_NMF_with_clusters6<- bind_rows(spoken_NMF_with_clusters6_plac, spoken_NMF_with_clusters6_drug)
```

some visualisations of spoken_NMF_with_clusters6
```{r}
unique(spoken_NMF_with_clusters6$filename)

# visualise example of substates
spoken_NMF_with_clusters6%>%
    # subset(idno == 30)%>%
  subset(filename == "./cut_spoken_sad_day1_p17.csv")%>%

  ggplot(aes(as.numeric(bin_frame), sp_avg))+

  geom_rect(aes(xmin = bin_frame, xmax = dplyr::lag(bin_frame), ymin = -Inf, ymax = Inf, fill = factor(cluster_across)), 
            alpha = .6)  +
    geom_line(size = 1)+
  geom_line(aes(y = comp_delta_avg),  size = 1, linetype = 2)+
  geom_line(aes(y = k3_comp1))+
  geom_line(aes(y = k3_comp2))+
  geom_line(aes(y = k3_comp3))+
     scale_fill_viridis_d(option = "magma")+
  theme_classic()

# visualise one profile for substates and facegrid variables


```

create cluster labels consistent with posed labels

```{r}
spoken_NMF_with_clusters6$cluster_label <- if_else(spoken_NMF_with_clusters6$cluster_across == 3, "relax",
                                                 if_else(spoken_NMF_with_clusters6$cluster_across == 1, "transition",
                                                         if_else(spoken_NMF_with_clusters6$cluster_across == 2,"sustain", NA)))


spoken_NMF_with_clusters6

spoken_NMF_with_clusters6


# peak speed displ

```


bind spoken and posed include timebins for new entropy calculation

```{r}

spoken_NMF_with_clusters6$comp1<- spoken_NMF_with_clusters6$k3_comp1
spoken_NMF_with_clusters6$comp2<- spoken_NMF_with_clusters6$k3_comp2
spoken_NMF_with_clusters6$comp3<- spoken_NMF_with_clusters6$k3_comp3

df_with_clusters6$comp1
spoken_NMF_with_clusters6
df_with_clusters6$
spoken_posed_nmf<- bind_rows(
  spoken_NMF_with_clusters6 [,c("filename","bin_frame", "subject", "cluster_label","expression","posed_spoken","drug.placebo","sp_avg","comp_delta_avg", "peak_speed", "max_displacement","comp1","comp2", "comp3")],
  df_with_clusters6[,c("filename","bin_frame","subject", "cluster_label","expression","posed_spoken","drug.placebo","sp_avg","comp_delta_avg", "peak_speed", "max_displacement","comp1","comp2", "comp3")])%>%
  group_by(filename, subject,bin_frame, cluster_label, expression, posed_spoken,
           drug.placebo)%>%
  summarise_if(is.numeric, mean, na.rm = T)

# do the entropy calculation again here

spoken_posed_nmf

# spoken_posed_nmf$cluster_label<- if_else(spoken_posed_nmf$cluster_label == "relax", "relaxed", spoken_posed_nmf$cluster_label)

unique(spoken_posed_nmf$cluster_label)
spoken_posed_nmf

```


```{r}

# recompute speed and displacement
spoken_posed_nmf%>%
  
  arrange(subject,filename,bin_frame)%>%
group_by(filename)%>%
   mutate(comp_avg = (comp1 +comp2+comp3)/3)%>%
  mutate(displacement = (comp_avg - lag(comp_avg, default = NA))) %>%
  mutate(time_difference = abs(bin_frame - lag(bin_frame, default = NA))) %>%
  mutate(speed = abs(displacement) / time_difference) %>%
  mutate(acceleration = ((speed - lag(speed, default = speed[1])) / time_difference))%>%
  subset(filename == "./cut_posed_angry_day1_p10.csv")%>%
  ggplot(aes(bin_frame, displacement, colour = "red"))+
  geom_line()+
  # geom_line(aes(y = comp2))+
  #  geom_line(aes(y = comp3))+
   # geom_line(aes(y = displacement), linetype = "dashed", colour = "red")+
   geom_line(aes(y = speed), linetype = "dashed", colour = "blue")+
   geom_line(aes(y = acceleration), linetype = "dashed", colour = "green")


# check peak and stuff

spoken_posed_nmf%>%
  
  arrange(subject,filename,bin_frame)%>%
group_by(filename)%>%
   mutate(comp_avg = (comp1 +comp2+comp3)/3)%>%
  mutate(displacement = (comp_avg - lag(comp_avg, default = NA))) %>%
  mutate(time_difference = abs(bin_frame - lag(bin_frame, default = NA))) %>%
  mutate(speed = abs(displacement) / time_difference) %>%
  mutate(acceleration = ((speed - lag(speed, default = speed[1])) / time_difference))%>%
  # subset(filename == "./cut_posed_angry_day1_p10.csv")%>%
  group_by(filename,cluster_label,posed_spoken,expression)%>%
  summarise_if(is.numeric, max, na.rm =T)%>%
  ggplot(aes(comp_avg, speed, colour =posed_spoken ))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_grid(cluster_label ~expression)


spoken_posed_nmf%>%
  
  arrange(subject,filename,bin_frame)%>%
group_by(filename)%>%
   mutate(comp_avg = (comp1 +comp2+comp3)/3)%>%
  mutate(displacement = abs(comp_avg - lag(comp_avg, default = NA))) %>%
  mutate(time_difference = abs(bin_frame - lag(bin_frame, default = NA))) %>%
  mutate(speed = abs(displacement) / time_difference) %>%
  mutate(acceleration = ((speed - lag(speed, default = speed[1])) / time_difference))%>%
  # subset(filename == "./cut_posed_angry_day1_p10.csv")%>%
  group_by(filename,cluster_label,posed_spoken,expression)%>%
  summarise_if(is.numeric, mean, na.rm =T)%>%
  ggplot(aes(expression, speed, colour =posed_spoken ))+
  stat_summary(geom = "pointrange")+
  geom_smooth(method = "lm")+
  facet_grid(~cluster_label)


spoken_posed_nmf<- spoken_posed_nmf%>%
  
  arrange(subject,filename,bin_frame)%>%
group_by(filename)%>%
   mutate(comp_avg_new = (comp1 +comp2+comp3)/3)%>%
  mutate(displacement = comp_avg - lag(comp_avg, default = NA)) %>%
  mutate(time_difference = abs(bin_frame - lag(bin_frame, default = NA))) %>%
  mutate(speed = abs(displacement) / time_difference) %>%
  mutate(acceleration = ((speed - lag(speed, default = speed[1])) / time_difference))

spoken_posed_nmf

```
 
 
 




```{r}
# spoken_posed_nmf_backup<- spoken_posed_nmf

spoken_posed_nmf
spoken_posed_nmf$cluster_label

# 
# colnames(spoken_posed_nmf)
# spoken_posed_nmf$sp_avg

# remove those without one of the sets


# no data for p9 as he refused to do the task and then day 1 data only for 

# Values to remove
tmp_subj_to_remove <- c(4, 9, 12, 26, 34, 35, 41, 42, 43)

# Dataframe filtering
spoken_posed_nmf_filtered <- spoken_posed_nmf[!spoken_posed_nmf$subject %in% tmp_subj_to_remove, ]%>%
    subset(expression!= "neutral")


spoken_posed_nmf_filtered_agg<- spoken_posed_nmf_filtered%>%
  group_by(subject, filename, expression,posed_spoken, drug.placebo, cluster_label)%>%
  summarise_if(is.numeric,mean, na.rm = T)


write_csv(spoken_posed_nmf_filtered_agg, "spoken_posed_nmf_filtered_agg.csv")


```


```{r}
spoken_posed_nmf_filtered$subject<- as.factor(spoken_posed_nmf_filtered$subject)



spoken_posed_nmf_filtered$sp_avg
spoken_posed_nmf_filtered_agg_bysubj<- spoken_posed_nmf_filtered%>%
  group_by(subject)%>%
  mutate(peak_speed_norm = normalize_0_1(peak_speed),
         entropy_clust_normlog = peak_speed/log2(entropy_clust))%>%
   group_by(subject, expression,posed_spoken,cluster_label)%>%
  summarise_if(is.numeric, mean, na.rm = T)
  


```


speed lmer
```{r}
library(lmerTest)
options(scipen = 999)
rslt_speedavg<-lmer(speed_avg~ expression*posed_spoken*cluster_label +(1|subject),
     REML =FALSE,
     data = spoken_posed_nmf_filtered_agg)

anova(rslt_speedavg)

# Main effects pairwise
emmeans::emmeans(rslt_speedavg, pairwise~cluster_label, adjust = "bonf")
emmeans::emmeans(rslt_speedavg, pairwise~expression, adjust = "bonf")
emmeans::emmeans(rslt_speedavg, pairwise~posed_spoken, adjust = "none")

# two way interactions
emmeans::emmeans(rslt_speedavg, pairwise~expression|cluster_label, adjust = "bonf")
emmeans::emmeans(rslt_speedavg, pairwise~expression|posed_spoken, adjust = "bonf")
emmeans::emmeans(rslt_speedavg, pairwise~cluster_label|posed_spoken, adjust = "bonf")

# three way interactions
emmeans::emmeans(rslt_speedavg, pairwise~expression|posed_spoken*cluster_label, adjust = "bonf")
emmeans::emmeans(rslt_speedavg, pairwise~posed_spoken|expression*cluster_label, adjust = "bonf")

anova(rslt_speedavg)
summary(rslt_speedavg)
report::report(rslt_speedavg)
report::report(anova(rslt_speedavg))
# tmp_emmeans<- emmeans::emmeans(rslt_speedavg, pairwise~expression|posed_spoken*cluster_label, adjust = "bonf")
# report::report(tmp_emmeans$contrasts)

saveRDS(rslt_speedavg,
        )


saveRDS(rslt_speedavg,
        file.path(paths$export_models, 
                  "mod_lmer_speed.rds"))
```






# entropy  - correct after fixing issue https://github.com/hcuve/halo_faces/issues/1

```{r}
library(data.table)
spoken_posed_nmf_filtered
spoken_posed_nmf_filtered_agg

library(dplyr)
library(data.table)  # for rleid
library(entropy)
library(GrpString)


```

```{r}

spoken_posed_nmf_entropy<- spoken_posed_nmf_filtered%>%
   
  group_by(filename) %>%
  mutate(entropy_clust = entropy::entropy(table(cluster_label))) %>%
    mutate(entropy_clust_norm =entropy_clust/(log2(3)))
  spoken_posed_nmf_entropy$trans_entropy
  
  
  spoken_posed_nmf_trans_entropy

```

  
  cor.test(  spoken_posed_nmf_entropy$entropy_clust_norm,
             spoken_posed_nmf_trans_entropy$trans_entropy_norm)


spoken_posed_nmf_entropy
# transition entropy correct
```{r}
max_transition_entropy <- function(n) {
  if (n > 1) {
    return(log2(n * n))
  } else {
    return(0)  # If there's only one state, entropy is zero
  }
}

log2(9)
log2(3)
max_transition_entropy(3)

# Function to normalize transition entropy
normalize_transition_entropy <- function(entropy, n_states) {
  return(entropy / max_transition_entropy(n_states))
}
n_states = 3


spoken_posed_nmf_trans_entropy<- spoken_posed_nmf_filtered%>%
    subset(expression!= "neutral")%>%
      group_by(filename, expression, subject,posed_spoken) %>%
      # keep just the initial of the cluster label, to avoid the actual words affecting entropy calculation
      summarise(cluster_label_combined = paste(substr(cluster_label, 1, 1), collapse = ""), .groups = 'drop')%>%
  group_by(filename)%>%
      # Compute transition entropy
      mutate(trans_entropy = GrpString::TransEntro(cluster_label_combined),
             trans_entropy_perclust_res = GrpString::TransEntropy(cluster_label_combined),
             trans_entropy_perclust_vec = GrpString::TransEntropy(cluster_label_combined)[,2]
         )%>%
  mutate(trans_entropy_norm=normalize_transition_entropy(trans_entropy, n_states))%>%
  ungroup()




```



lmer entropy


entropy

```{r}

# stationary entropy
spoken_posed_nmf_entropy #stat
spoken_posed_nmf_trans_entropy

spoken_posed_nmf_entropy$subject <- as.factor(spoken_posed_nmf_entropy$subject)

rslt_stat_entropy<-lmer(log1p(entropy_clust_norm+.01)~ expression*posed_spoken +(1+posed_spoken+expression|subject),
     REML =FALSE,
     data = spoken_posed_nmf_entropy%>%
       subset(expression!= "neutral"))

qqnorm(resid(rslt_stat_entropy))
summary(rslt_stat_entropy)
anova(rslt_stat_entropy)


summary(rslt_stat_entropy)



emmeans::emmeans(rslt_stat_entropy, pairwise~expression, adjust = "bonf")
emmeans::emmeans(rslt_stat_entropy, pairwise~posed_spoken)

emmeans::emmeans(rslt_stat_entropy, pairwise~expression|posed_spoken, adjust = "holm")
# install.packages("report")
options(scipen = 999)
report::report(rslt_stat_entropy)
report::report(anova(rslt_stat_entropy))

rslt_stat_entropy


# rslt_stat_entropy@frame you can find the data here
saveRDS(rslt_stat_entropy,
        file.path(paths$export_models, 
                  "mod_lmer_stat_entropy.rds"))
```


We fitted a linear mixed model (estimated using ML and nloptwrap optimizer) to predict entropy_clust_norm with expression and posed_spoken (formula:
sqrt(entropy_clust_norm + 0.1) ~ expression * posed_spoken). The model included posed_spoken as random effects (formula: ~1 + posed_spoken | subject).
The model's total explanatory power is substantial (conditional R2 = 0.44) and the part related to the fixed effects alone (marginal R2) is of 0.27.
The model's intercept, corresponding to expression = angry and posed_spoken = posed, is at 0.81 (95% CI [0.81, 0.82], t(40739) = 276.12, p < .001).
Within this model:

  - The effect of expression [happy] is statistically significant and negative (beta = -3.39e-03, 95% CI [-4.46e-03, -2.32e-03], t(40739) = -6.19, p <
.001; Std. beta = -0.02, 95% CI [-0.03, -0.02])
  - The effect of expression [sad] is statistically significant and negative (beta = -4.48e-03, 95% CI [-5.55e-03, -3.40e-03], t(40739) = -8.18, p <
.001; Std. beta = -0.02, 95% CI [-0.03, -0.02])
  - The effect of posed spoken [spoken] is statistically significant and positive (beta = 0.04, 95% CI [0.03, 0.05], t(40739) = 10.22, p < .001; Std.
beta = 0.23, 95% CI [0.18, 0.28])
  - The effect of expression [happy] × posed spoken [spoken] is statistically significant and negative (beta = -7.82e-03, 95% CI [-9.34e-03, -6.30e-03],
t(40739) = -10.09, p < .001; Std. beta = -0.03, 95% CI [-0.04, -0.02])
  - The effect of expression [sad] × posed spoken [spoken] is statistically significant and positive (beta = 9.50e-03, 95% CI [7.98e-03, 0.01], t(40739)
= 12.26, p < .001; Std. beta = 0.05, 95% CI [0.05, 0.06])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were
computed using a Wald t-distribution approximation.

The ANOVA suggests that:

  - The main effect of expression is statistically significant and large (F(2) = 3.29, p = 0.049; Eta2 (partial) = 0.16, 95% CI [2.30e-04, 1.00])
  - The main effect of posed_spoken is statistically significant and large (F(1) = 110.07, p < .001; Eta2 (partial) = 0.76, 95% CI [0.64, 1.00])
  - The interaction between expression and posed_spoken is statistically significant and small (F(2) = 315.65, p < .001; Eta2 (partial) = 0.02, 95% CI
[0.01, 1.00])

transition entropy (main for the paper) lmer

```{r}
rslt_trans_entropy<-lmer(log1p(trans_entropy_norm+.01)~ expression*posed_spoken +(1+posed_spoken|subject),
     REML =FALSE,
     data = spoken_posed_nmf_trans_entropy%>%
       subset(expression!= "neutral"))


plot(rslt_trans_entropy)
qqnorm(resid(rslt_trans_entropy))
qqline(resid(rslt_trans_entropy))
summary(rslt_trans_entropy)
anova(rslt_trans_entropy)


summary(rslt_trans_entropy)



emmeans::emmeans(rslt_trans_entropy, pairwise~expression, adjust = "bonf")
emmeans::emmeans(rslt_trans_entropy, pairwise~posed_spoken)

emmeans::emmeans(rslt_trans_entropy, pairwise~expression|posed_spoken, adjust = "holm")
# install.packages("report")
report::report(rslt_trans_entropy)
report::report(anova(rslt_trans_entropy))




saveRDS(rslt_trans_entropy,
        file.path(paths$export_models, 
                  "mod_lmer_trans_entropy.rds"))
```

We fitted a linear mixed model (estimated using ML and nloptwrap optimizer) to predict trans_entropy_norm with expression and posed_spoken (formula:
log1p(trans_entropy_norm + 0.01) ~ expression * posed_spoken). The model included posed_spoken as random effects (formula: ~1 + posed_spoken |
subject). The model's total explanatory power is substantial (conditional R2 = 0.63) and the part related to the fixed effects alone (marginal R2) is
of 0.57. The model's intercept, corresponding to expression = angry and posed_spoken = posed, is at 0.43 (95% CI [0.42, 0.44], t(398) = 65.64, p <
.001). Within this model:

  - The effect of expression [happy] is statistically non-significant and negative (beta = -9.76e-03, 95% CI [-0.03, 5.49e-03], t(398) = -1.26, p =
0.209; Std. beta = -0.03, 95% CI [-0.08, 0.01])
  - The effect of expression [sad] is statistically significant and positive (beta = 0.02, 95% CI [1.81e-03, 0.03], t(398) = 2.20, p = 0.028; Std. beta =
0.05, 95% CI [8.09e-03, 0.10])
  - The effect of posed spoken [spoken] is statistically significant and positive (beta = 0.11, 95% CI [0.10, 0.13], t(398) = 13.53, p < .001; Std. beta
= 0.33, 95% CI [0.28, 0.38])
  - The effect of expression [happy] × posed spoken [spoken] is statistically non-significant and negative (beta = -2.51e-03, 95% CI [-0.02, 0.02],
t(398) = -0.23, p = 0.819; Std. beta = 2.98e-03, 95% CI [-0.06, 0.07])
  - The effect of expression [sad] × posed spoken [spoken] is statistically non-significant and negative (beta = -2.10e-03, 95% CI [-0.02, 0.02], t(398)
= -0.19, p = 0.848; Std. beta = -0.01, 95% CI [-0.08, 0.05])






The ANOVA suggests that:

  - The main effect of expression is statistically significant and medium (F(2) = 12.27, p < .001; Eta2 (partial) = 0.07, 95% CI [0.03, 1.00])
  
  
  - The main effect of posed_spoken is statistically significant and large 
  (F(1) = 429.10, p < .001; Eta2 (partial) = 0.93, 95% CI [0.89, 1.00])
  
  - The interaction between expression and posed_spoken is statistically not significant and very small (F(2) = 0.03, p = 0.970; Eta2 (partial) =
1.77e-04, 95% CI [0.00, 1.00])

Standardized parameters were obtained by fitting the model on a standardized version of the dataset. 95% Confidence Intervals (CIs) and p-values were
computed using a Wald t-distribution approximation.



$emmeans
 expression emmean      SE  df lower.CL upper.CL
 angry       0.486 0.00478 108    0.476    0.495
 happy       0.475 0.00478 108    0.465    0.484
 sad         0.502 0.00478 108    0.492    0.511

Results are averaged over the levels of: posed_spoken 
Degrees-of-freedom method: kenward-roger 
Results are given on the log1p(mu + 0.01) (not the response) scale. 
Confidence level used: 0.95 

$contrasts
 contrast      estimate      SE  df t.ratio p.value
 angry - happy    0.011 0.00552 344   1.996  0.1401
 angry - sad     -0.016 0.00552 344  -2.901  0.0119
 happy - sad     -0.027 0.00552 344  -4.897  <.0001

Results are averaged over the levels of: posed_spoken 
Note: contrasts are still on the log1p(mu + 0.01) scale 
Degrees-of-freedom method: kenward-roger 
P value adjustment: bonferroni method for 3 tests 
eigen(rslt_trans_entropy)

  
summary(rslt_trans_entropy)



