---
title: "preprocess_halofacestudy"
author: "Hélio Clemente Cuve"
date: '2022-06-06'
output: html_document
---


#experimental conditions
3 expression conditions (posed, spoken and spoken2)
3 emotion categories (happy, angry, sad)
Spoken has 4 emotions (H. A. S and Neutral)
Spoken 2 is the additional for all but was never analysed before

posed = expression only (term used in the paper)
spoken1 = emotive speech (term used in the paper)

# labels
each video is labeled with the day (1 or 2) - see unblinding_BMY_drugday
- this is needed to group the drug vs placebo correctly

notes
- some participants we only have day 1 data 

no data for p9 as he refused to do the task and then day 1 data only for 4, 12, 26, 34, 35, 41, 42 and 43 

Setup preferences
```{r}
# set working directory
# setwd("...HaloStudy/Data/openFaceProcessed")
# setwd(".../RE-CUT_VIDEOS_23.05.22-20220523T152230Z-001/RE-CUT_VIDEOS_23.05.22/openFaceProcessed")

# load libraries
library(tidyverse)
library(data.table)
library(readr)
library(stringr)

```

Load 'csv' OpenFace File and Bind into One Large DataFrame

```{r}
# Function to read and append filename to dataframe
read_plus <- function(flnm) {
  data.table::fread(flnm) %>% 
    mutate(filename = flnm)
}

# Read files and bind them into one dataframe
df_OF_output <- list.files(pattern = "*.csv", full.names = TRUE) %>% 
  map_df(~read_plus(.))

# Quick inspection of the data
df_OF_output
dim(df_OF_output)
colnames(df_OF_output)
table(is.na(df_OF_output$filename))

# View first 10 and last 5 rows
head(df_OF_output, n = 10)
tail(df_OF_output, n = 5)


```

# Data Reduction, Merging, and Visualisation
1. Reduce the dataset to the needed Action Units (AUs)
2. Merge with unblinding data
3. Quick visualisations

```{r}
colnames(df_OF_output)
df_OF_output_AUs_land <- df_OF_output[, c(1:5, 300:435, 680:715)]

# Raw data for OSF
dta_pos_spk1_lan_aus <- df_OF_output_AUs_land %>%
  subset(!grepl("spoken2", filename))

setwd(".../My Drive/2022 - University Of Birmingham/HaloStudy/Data/OSF")

# Split the data based on the levels of filename and save each as a CSV
dta_pos_spk1_lan_aus$filename1 <- str_remove(dta_pos_spk1_lan_aus$filename, "./")


dta_pos_spk1_lan_aus$subject<- sub(".*_p", "", stringr::str_match(dta_pos_spk1_lan_aus$filename,                                                        "p\\s*(.*?)\\s*.csv")[,2])


dta_pos_spk1_lan_aus$subject<- as.factor(as.character(dta_pos_spk1_lan_aus$subject))

# testday

dta_pos_spk1_lan_aus$test.day<- stringr::str_match(dta_pos_spk1_lan_aus$filename, "_day\\s*(.*?)\\s*_p")[,2]

# expression
dta_pos_spk1_lan_aus$expression <-sapply(strsplit(dta_pos_spk1_lan_aus$filename, "_"), function(x) x[3])

dta_pos_spk1_lan_aus_1<- dta_pos_spk1_lan_aus%>%
  subset(expression!= "neutral")%>%
  mutate(expression = as.character(expression))

unique(dta_pos_spk1_lan_aus_1$expression)
colnames(dta_pos_spk1_lan_aus_1)

split_data <- split(dta_pos_spk1_lan_aus_1, dta_pos_spk1_lan_aus_1$filename1)

colnames(split_data$cut_posed_angry_day1_p1.csv)

# Iterate over each element in the split data list and save it to a CSV file
lapply(names(split_data), function(name) {
  
  # file_name <- paste0(paths$raw_data, name) #name already has csv in the name
  file_name <- paste0(name)
  write.csv(split_data[[name]], 
            file_name, 
            row.names = FALSE)
})

colnames(df_OF_output_AUs_land)

# df_OF_output_AUs_land
# saveRDS(df_OF_output_AUs_land,"df_OF_output_AUs_land_unblind.rds") # to do link where this rds is exported from
```

colnames(split_data$cut_posed_angry_day1_p1.csv)
[1] "frame"      "face_id"    "timestamp"  "confidence" "success"   
  [6] "x_0"        "x_1"        "x_2"        "x_3"        "x_4"       
 [11] "x_5"        "x_6"        "x_7"        "x_8"        "x_9"       
 [16] "x_10"       "x_11"       "x_12"       "x_13"       "x_14"      
 [21] "x_15"       "x_16"       "x_17"       "x_18"       "x_19"      
 [26] "x_20"       "x_21"       "x_22"       "x_23"       "x_24"      
 [31] "x_25"       "x_26"       "x_27"       "x_28"       "x_29"      
 [36] "x_30"       "x_31"       "x_32"       "x_33"       "x_34"      
 [41] "x_35"       "x_36"       "x_37"       "x_38"       "x_39"      
 [46] "x_40"       "x_41"       "x_42"       "x_43"       "x_44"      
 [51] "x_45"       "x_46"       "x_47"       "x_48"       "x_49"      
 [56] "x_50"       "x_51"       "x_52"       "x_53"       "x_54"      
 [61] "x_55"       "x_56"       "x_57"       "x_58"       "x_59"      
 [66] "x_60"       "x_61"       "x_62"       "x_63"       "x_64"      
 [71] "x_65"       "x_66"       "x_67"       "y_0"        "y_1"       
 [76] "y_2"        "y_3"        "y_4"        "y_5"        "y_6"       
 [81] "y_7"        "y_8"        "y_9"        "y_10"       "y_11"      
 [86] "y_12"       "y_13"       "y_14"       "y_15"       "y_16"      
 [91] "y_17"       "y_18"       "y_19"       "y_20"       "y_21"      
 [96] "y_22"       "y_23"       "y_24"       "y_25"       "y_26"      
[101] "y_27"       "y_28"       "y_29"       "y_30"       "y_31"      
[106] "y_32"       "y_33"       "y_34"       "y_35"       "y_36"      
[111] "y_37"       "y_38"       "y_39"       "y_40"       "y_41"      
[116] "y_42"       "y_43"       "y_44"       "y_45"       "y_46"      
[121] "y_47"       "y_48"       "y_49"       "y_50"       "y_51"      
[126] "y_52"       "y_53"       "y_54"       "y_55"       "y_56"      
[131] "y_57"       "y_58"       "y_59"       "y_60"       "y_61"      
[136] "y_62"       "y_63"       "y_64"       "y_65"       "y_66"      
[141] "y_67"       "AU01_r"     "AU02_r"     "AU04_r"     "AU05_r"    
[146] "AU06_r"     "AU07_r"     "AU09_r"     "AU10_r"     "AU12_r"    
[151] "AU14_r"     "AU15_r"     "AU17_r"     "AU20_r"     "AU23_r"    
[156] "AU25_r"     "AU26_r"     "AU45_r"     "AU01_c"     "AU02_c"    
[161] "AU04_c"     "AU05_c"     "AU06_c"     "AU07_c"     "AU09_c"    
[166] "AU10_c"     "AU12_c"     "AU14_c"     "AU15_c"     "AU17_c"    
[171] "AU20_c"     "AU23_c"     "AU25_c"     "AU26_c"     "AU28_c"    
[176] "AU45_c"     "filename"   "filename1"  "subject"    "test.day"  
[181] "expression"

# rename action units based on a dictionary I created for the NMF tutorial

```{r}
# Load AU code dictionaries
AU_OF_codes_merged <- readr::read_csv("AU_OF_codes_merged.csv")
AU_OF_transposed <- read_csv("AU_OF_transposed.csv")
AU_Wiki <- read_csv("AU_Wiki.csv")

# AU weights
AU_OF_codes_merged$AU_Code_OF_r <- paste0(AU_OF_codes_merged$AU_Code_OF, "_", AU_OF_codes_merged$FACS_name)

# Codes for AU presence
AU_OF_codes_merged$AU_Code_OF_c <- paste0("AU", AU_OF_codes_merged$AU_CODE, "_c_", AU_OF_codes_merged$FACS_name)

# Rename AUs based on appended names and codes
cnames1 <- colnames(df_OF_output_AUs_land)
cname2 <- c(cnames1[1:141], t(AU_OF_codes_merged$AU_Code_OF_r), t(AU_OF_codes_merged$AU_Code_OF_c), "filename")
df_OF_output_AUs$AU28_c <- NULL  # Exclude AU28_c for now
names(df_OF_output_AUs) <- cname2


colnames(df_OF_output_AUs_land)

# quick check
unique(df_OF_output_AUs$filename)
df_OF_output_AUs%>%
  ggplot(aes(AU01_r))+
  geom_histogram()

df_OF_output_AUs%>%
  ggplot(aes(AU01_c))+
  geom_histogram()


colnames(df_OF_output_AUs)
```

Merge Unblinding Data
```{r}

library(readxl)
Unblinding_BMI_drugday <- read_excel("Unblinding_BMI_drugday.xlsx")
View(Unblinding_BMI_drugday)

df_OF_output_AUs$subject<- sub(".*_p", "", str_match(df_OF_output_AUs$filename, "p\\s*(.*?)\\s*.csv")[,2])

# testday

df_OF_output_AUs$test.day<- str_match(df_OF_output_AUs$filename, "_day\\s*(.*?)\\s*_p")[,2]


unique(df_OF_output_AUs$subject)
# how many participants do we expect = 43, checks out
Unblinding_BMI_drugday

names(Unblinding_BMI_drugday)<- make.names(names(Unblinding_BMI_drugday))
Unblinding_BMI_drugday

# merge
colnames(df_OF_output_AUs)
df_OF_output_AUs$subject<- as.factor(as.character( df_OF_output_AUs$subject))
Unblinding_BMI_drugday$subject<-as.factor(as.character(Unblinding_BMI_drugday$subject))

# expected rows()
# nrow(df_OF_output_AUs)  =131476
df_OF_output_AUs_unblind<- left_join(df_OF_output_AUs, Unblinding_BMI_drugday)

# if test.day == drugday  = "halo, vs placebo
colnames(df_OF_output_AUs_unblind)

df_OF_output_AUs_unblind$drug.placebo <- if_else(df_OF_output_AUs_unblind$test.day == df_OF_output_AUs_unblind$drug.day, "drug", "placebo")

# add emotion
# return word between 3rd and 4rth underscore df$values = sapply(strsplit(df$V1, "_"), function(x) x[3])
unique( sapply(strsplit(df_OF_output_AUs_unblind$filename, "_"), function(x) x[3]))

df_OF_output_AUs_unblind$expression <-sapply(strsplit(df_OF_output_AUs_unblind$filename, "_"), function(x) x[3])
# 2nd and 3rd undrcore
# unique( sapply(strsplit(df_OF_output_AUs_unblind$filename, "_"), function(x) x[2]))
df_OF_output_AUs_unblind$posed.spoken<- sapply(strsplit(df_OF_output_AUs_unblind$filename, "_"), function(x) x[2])


unique(df_OF_output_AUs_unblind$posed.spoken)

```

Do some quick visualizations for time, etc

```{r}
# Distribution of duration for test day 1
df_OF_output_AUs_unblind %>%
  subset(test.day == 1) %>%
  group_by(filename, test.day, expression, posed.spoken) %>%
  summarise_at(c("timestamp", "frame"), max, na.rm = TRUE) %>%
  ggplot(aes(frame)) +
  geom_histogram() +
  theme_classic() +
  facet_grid(expression ~ posed.spoken)

# Distribution of duration for test day 2
df_OF_output_AUs_unblind %>%
  subset(test.day == 2) %>%
  group_by(filename, test.day, expression, posed.spoken) %>%
  summarise_at(c("timestamp", "frame"), max, na.rm = TRUE) %>%
  ggplot(aes(timestamp)) +
  geom_histogram() +
  theme_classic() +
  facet_grid(expression ~ posed.spoken)
```

Visualise heatmaps for action units

```{r}
# store weights only
colnames(df_OF_output_AUs_unblind)
df_OF_output_AUsW_unblind<- df_OF_output_AUs_unblind[,c(1:5,40:49,6:22)]

df_OF_output_AUsW_unblind%>%
  gather(AUs, AU.coefs, -(1:15))%>%
  group_by(timestamp,frame, AUs, expression,drug.placebo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame), AUs, fill = AU.coefs))+
    geom_tile()+
    facet_grid(drug.placebo~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())


df_OF_output_AUsW_unblind%>%
  gather(AUs, AU.coefs, -(1:15))%>%
  group_by(timestamp,frame, AUs,posed.spoken,drug.placebo)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame), AUs, fill = AU.coefs))+
    geom_tile()+
    facet_grid(posed.spoken~drug.placebo)+
    theme_classic()+
  theme(axis.text.y = element_blank())+

  scale_fill_viridis_c(option = "magma")
  
  # check emotion differences
  df_OF_output_AUsW_unblind%>%
  gather(AUs, AU.coefs, -(1:15))%>%
  group_by(timestamp,frame, AUs,expression,posed.spoken)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame), AUs, fill = AU.coefs))+
    geom_tile()+
    facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  # scale_fill_viridis_c(option = "plasma")
  # scale_fill_viridis_c()
  scale_fill_viridis_c(option = "magma")


```


Binify Based on Moving Average for Comparable Bins

start by separating just emotion

```{r}
df_OF_output_AUsW_unblind_posed<- subset(df_OF_output_AUsW_unblind, 
                                         posed.spoken == "posed")

df_OF_output_AUsW_unblind_posed

df_OF_output_AUsW_unblind_posed%>%
  subset(test.day ==1)%>%
  group_by(filename,test.day, expression,posed.spoken)%>%
  summarise_at(c("timestamp", "frame"), max, na.rm = T)%>%
  ggplot(aes(frame))+
  geom_histogram()+
  theme_classic()+
  facet_grid(expression~posed.spoken)


 # = 100 bins
# average is 250
# 220/100
colnames(df_OF_output_AUsW_unblind_posed)
df_OF_output_AUsW_unblind_posed
?cut




df_OF_output_AUsW_unblind_posed_binned <-df_OF_output_AUsW_unblind_posed%>%
  group_by(filename, subject,test.day, age, gender,BMI,drug.day, drug.placebo, expression )%>%
  mutate(bin_frame = cut(frame, 100, labels = FALSE))%>%
  group_by(filename,bin_frame,subject,test.day, age, gender,BMI,drug.day, drug.placebo, expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
colnames(df_OF_output_AUsW_unblind_posed_binned) 
 
df_OF_output_AUsW_unblind_posed_binned$filename
# "./cut_posed_angry_day1_p1.csv"
colnames(df_OF_output_AUsW_unblind_posed_binned)
df_OF_output_AUsW_unblind_posed_binned%>%
  subset(filename == "./cut_posed_angry_day1_p1.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(bin_frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(bin_frame), AUs, fill = AU.coefs))+
    geom_tile()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")

# unbinned
df_OF_output_AUsW_unblind_posed%>%
  subset(filename == "./cut_posed_angry_day1_p1.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame), AUs, fill = AU.coefs))+
    geom_tile()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


# similar
# let's check lines

df_OF_output_AUsW_unblind_posed_binned%>%
  subset(filename == "./cut_posed_angry_day1_p1.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(bin_frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(bin_frame), AU.coefs, group = AUs))+
    geom_line()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


df_OF_output_AUsW_unblind_posed%>%
  subset(filename == "./cut_posed_angry_day1_p1.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame),AU.coefs, group = AUs))+
 geom_line()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


df_OF_output_AUsW_unblind_posed_binned%>%
  subset(subject == 1)%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(bin_frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(bin_frame), AU.coefs, group = AUs))+
   geom_smooth(se = F)+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


df_OF_output_AUsW_unblind_posed%>%
  subset(subject == 1)%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame),AU.coefs, group = AUs))+
  geom_smooth(se = F)+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


table(df_OF_output_AUsW_unblind_posed_binned$bin_frame)

```


Pre-Process speech data
note we need only spoken 1 data

```{r}
unique(df_OF_output_AUsW_unblind$posed.spoken)
df_OF_output_AUsW_unblind_spoken1_2<- subset(df_OF_output_AUsW_unblind, 
                                         posed.spoken != "posed")

df_OF_output_AUsW_unblind_spoken1_2

# check how variable is duration
df_OF_output_AUsW_unblind_spoken1_2%>%
  subset(test.day ==1)%>%
  group_by(filename,test.day, expression,posed.spoken)%>%
  summarise_at(c("timestamp", "frame"), max, na.rm = T)%>%
  ggplot(aes(frame))+
  geom_histogram()+
  theme_classic()+
  facet_grid(expression~posed.spoken)

```


```{r}
df_OF_output_AUsW_unblind_spoken1_2
 # = 100 bins
# average is 250
# 220/100

df_OF_output_AUsW_unblind_spoken1_2

# ?cut
colnames(df_OF_output_AUsW_unblind_spoken1_2)

# create bins

df_OF_output_AUsW_unblind_spoken1_2$posed.spoken
df_OF_output_AUsW_unblind_spoken1_2_binned <-df_OF_output_AUsW_unblind_spoken1_2%>%
  group_by(filename, subject,test.day, age, gender,BMI,drug.day, drug.placebo, expression,posed.spoken )%>%
  mutate(bin_frame = cut(frame, 100, labels = FALSE))%>%
  group_by(filename,bin_frame,subject,test.day, age, gender,BMI,drug.day, drug.placebo, expression,posed.spoken)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
 
 

df_OF_output_AUsW_unblind_spoken1_2_binned
df_OF_output_AUsW_unblind_spoken1_2_binned

df_OF_output_AUsW_unblind_spoken1_2_binned$filename

# "./cut_spoken_angry_day1_p19.csv"
colnames(df_OF_output_AUsW_unblind_spoken1_2_binned)

# quick visualisation of how the bins would be
df_OF_output_AUsW_unblind_spoken1_2_binned%>%
  subset(filename == "./cut_spoken_angry_day1_p19.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(bin_frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(bin_frame), AUs, fill = AU.coefs))+
    geom_tile()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


# unbined for comparison
df_OF_output_AUsW_unblind_spoken1_2%>%
  subset(filename == "./cut_spoken_angry_day1_p19.csv")%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame), AUs, fill = AU.coefs))+
    geom_tile()+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


# very similar, so we can be confident that the binification is not screwing things up
```



compare smoothed lines for binned vs unbinned spoken data
```{r}
df_OF_output_AUsW_unblind_spoken1_2_binned%>%
  # subset(filename == "./cut_spoken_angry_day1_p19.csv")%>%
  subset(subject == 1)%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(bin_frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(bin_frame), AU.coefs, group = AUs))+
   geom_smooth(se = F)+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")



df_OF_output_AUsW_unblind_spoken1_2%>%
  subset(subject == 1)%>%
    gather(AUs, AU.coefs, -(1:15))%>%
  group_by(frame, AUs,expression)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
  # subset(drug.placebo == "drug")%>%
    ggplot(aes(as.numeric(frame),AU.coefs, group = AUs))+
  geom_smooth(se = F)+
    # facet_grid(posed.spoken~expression)+
    theme_classic()+
  theme(axis.text.y = element_blank())+
  scale_fill_viridis_c(option = "magma")


table(ç$bin_frame)

unique(df_OF_output_AUsW_unblind_spoken1_2$posed.spoken)
```


Demographics

```{r}
Unblinding_BMI_drugday%>%
  group_by(gender)%>%
  summarize_if(is.numeric, mean, na.rm = T)

mean(Unblinding_BMI_drugday$age)

table(Unblinding_BMI_drugday$gender)

```


Main outputs

df_OF_output_AUsW_unblind_posed_binned
df_OF_output_AUsW_unblind_spoken1_2



