---
title: "Untitled"
author: "Helio"
date: "2025-03-27"
output: html_document
---




mod_pls_nmfaus_posed
saveRDS(mod_pls_nmfaus_posed,
          file.path(paths$au_to_land, 
                    "mod_pos_pls_nmfaus_posed.rds"))


allcomps_pred1_land_nmfaus <- predict(mod_pls_nmfaus_posed,allcomps_pred1_norm , ncomp=17)


mod_pls_spoken

predicted_spoken1_landmarks_sok_pls <- predict(mod_pls_spoken, dta_reconstr_spoken1_combined_agg1_0, ncomp = 17)


```{r}
mod_pls_nmfaus_posed
mod_pls_spoken
require(NMF)

#we need the test set projection for posed and spoken

mod_rslt_nmfposed_test_projection <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Models/OSF/mod_rslt_nmfposed_test_projection.Rds")
 mod_rslt_spk_nmf_test <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Models/OSF/mod_rslt_spk_nmf_test.rds")

dta_pos_nmf_proj<- NMF::fitted(mod_rslt_nmfposed_test_projection)

dta_spk_nmf_proj<- NMF::fitted(mod_rslt_spk_nmf_test)




# dta_expr_only_4nmf_test <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Processeddata/dta_expr_only_4nmf_test.rds")
dta_spk_4nmf_test <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Processeddata/dta_spk_4nmf_test.rds")
mod_rslt_spk_nmf_test
nrow(dta_pos_nmf_proj)
nrow(dta_expr_only_4nmf_test)
nrow(dta_spk_nmf_proj)
nrow(dta_spk_4nmf_test)


# transform into dataframes
dta_spk_nmf_proj

dta_pos_nmf_proj<- as.data.frame(dta_pos_nmf_proj)
dta_spk_nmf_proj<- as.data.frame(dta_spk_nmf_proj)

# add meatada

dta_pos_nmf_proj$filename<- dta_expr_only_4nmf_test$filename
dta_pos_nmf_proj$bin_frame<- dta_expr_only_4nmf_test$bin_frame
dta_pos_nmf_proj$subject<- dta_expr_only_4nmf_test$subject
dta_pos_nmf_proj$test.day<- dta_expr_only_4nmf_test$test.day
dta_pos_nmf_proj$expression<- dta_expr_only_4nmf_test$expression


dta_spk_nmf_proj$filename<- dta_spk_4nmf_test$filename
dta_spk_nmf_proj$bin_frame<- dta_spk_4nmf_test$bin_frame
dta_spk_nmf_proj$subject<- dta_spk_4nmf_test$subject
dta_spk_nmf_proj$test.day<- dta_spk_4nmf_test$test.day
dta_spk_nmf_proj$expression<- dta_spk_4nmf_test$expression

# spk


```

# now predict landmark from AUS


```{r}

predict(mod_pls_nmfaus_posed,allcomps_pred1_norm , ncomp=17)

au_cols<- colnames(allcomps_pred1)[2:18]

normalize_rowwise <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

require(tidyverse)

dta_pos_nmf_proj_norm<-  dta_pos_nmf_proj%>%
  group_by(filename)%>%
  mutate(across(all_of(au_cols), 
                ~ if_else(row_number() %in% rows_to_normalize, 
                          normalize_rowwise(as.numeric(.)), as.numeric(.))))

dta_spk_nmf_proj_norm<-  dta_spk_nmf_proj%>%
  group_by(filename)%>%
  mutate(across(all_of(au_cols), 
                ~ if_else(row_number() %in% rows_to_normalize, 
                          normalize_rowwise(as.numeric(.)), as.numeric(.))))






```

actually what we want is to use the nm model to project on the whole test set


```{r}
mod_rslt_nmfposed_train <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Models/OSF/mod_rslt_nmfposed_train.Rds")


dta_4_plfs_orig <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Processeddata/dta_4_plfs_orig.Rds")


dta_4_plfs_orig_pos<- dta_4_plfs_orig%>%
  subset(posed.spoken == "posed")

dta_4_plfs_orig_spk<- dta_4_plfs_orig%>%
  subset(posed.spoken != "posed")

mod_rslt_nmfposed_train@fit@W #time

colnames(dta_4_plfs_orig_pos)

mod_4_plfs_pos_nmf_proj<- nmf(dta_4_plfs_orig_pos[,139:155]+.0001, 
         H = mod_rslt_nmfposed_train@fit@H, # we actually want to use H
         rank= 3,
          seed=123456,
         # method = "lee", 
         nrun = 1)


 mod_rslt_spk_nmf_train <- readRDS("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedData/Models/OSF/mod_rslt_spk_nmf_train.rds")
 
mod_4_plfs_spk_nmf_proj<- nmf(dta_4_plfs_orig_spk[,139:155]+.0001, 
         H = mod_rslt_spk_nmf_train@fit@H, # we actually want to use H
         rank= 3,
          seed=123456,
         # method = "lee", 
         nrun = 1)


dta_4_plfs_pos_nmf_proj<- fitted(mod_4_plfs_pos_nmf_proj)
dta_4_plfs_spk_nmf_proj<- fitted(mod_4_plfs_spk_nmf_proj)

dta_4_plfs_pos_nmf_proj<- as.data.frame(dta_4_plfs_pos_nmf_proj)

dta_4_plfs_spk_nmf_proj<- as.data.frame(dta_4_plfs_spk_nmf_proj)



dta_4_plfs_pos_nmf_proj$filename<- dta_4_plfs_orig_pos$filename
# dta_4_plfs_pos_nmf_proj$bin_frame<- dta_4_plfs_orig_pos$bin_frame
dta_4_plfs_pos_nmf_proj$subject<- dta_4_plfs_orig_pos$subject
dta_4_plfs_pos_nmf_proj$test.day<- dta_4_plfs_orig_pos$test.day
dta_4_plfs_pos_nmf_proj$expression<- dta_4_plfs_orig_pos$expression
dta_4_plfs_pos_nmf_proj$frame<- dta_4_plfs_orig_pos$frame


dta_4_plfs_spk_nmf_proj$filename<- dta_4_plfs_orig_spk$filename
# dta_4_plfs_spk_nmf_proj$bin_frame<- dta_4_plfs_orig_spk$bin_frame
dta_4_plfs_spk_nmf_proj$subject<- dta_4_plfs_orig_spk$subject
dta_4_plfs_spk_nmf_proj$test.day<- dta_4_plfs_orig_spk$test.day
dta_4_plfs_spk_nmf_proj$expression<- dta_4_plfs_orig_spk$expression
dta_4_plfs_spk_nmf_proj$frame<- dta_4_plfs_orig_spk$frame

dta_4_plfs_pos_nmf_proj$frame
dta_4_plfs_spk_nmf_proj

dta_4_plfs_spk_nmf_proj
```
Now predict landmarks fro aus

```{r}

dta_4_plfs_pos_nmf_proj
dta_4_plfs_spk_nmf_proj


# range(dta_4_plfs_pos_nmf_proj$AU26_r_Jaw_drop)

dta_4_plfs_pos_nmf_proj_norm<-  dta_4_plfs_pos_nmf_proj%>%
    group_by(filename)%>%
    mutate(across(.cols = all_of(col_aus), 
                .fns = normalize_between_0_and_1))%>%
  ungroup()

range(dta_4_plfs_pos_nmf_proj_norm[,cols_to_normalize])

dta_4_plfs_spk_nmf_proj_norm<-  dta_4_plfs_spk_nmf_proj%>%
 group_by(filename)%>%
    mutate(across(.cols = all_of(col_aus), 
                .fns = normalize_between_0_and_1))%>%
  ungroup()

range(dta_4_plfs_spk_nmf_proj_norm[,cols_to_normalize])

range(dta_4_plfs_spk_nmf_proj_norm[,cols_to_normalize])

# predict landmarks
library(pls)

options(scipen = 999)
colnames(dta_4_plfs_pos_nmf_proj_norm)

dta_4_plfs_pos_nmf_proj_norm

# dta_4_plfs_pos_nmf_land_proj<- predict(mod_pls,dta_4_plfs_pos_nmf_proj_norm , ncomp=17)

# FOR THIS PURPOSE WE NEED TO TRAIN THE PLS CLASSIFIER SEPARATELLY FOR EACH EMOTION

dta_4_plfs_pos_nmf_land_proj<- predict(mod_pls_test,dta_4_plfs_pos_nmf_proj_norm , ncomp=17)

dta_4_plfs_pos_nmf_proj_norm_an<- dta_4_plfs_pos_nmf_proj_norm%>%
  subset(expression == "angry")

dta_4_plfs_pos_nmf_proj_norm_sad<- dta_4_plfs_pos_nmf_proj_norm%>%
  subset(expression == "sad")


dta_4_plfs_pos_nmf_proj_norm_hap<- dta_4_plfs_pos_nmf_proj_norm%>%
  subset(expression == "happy")
# posed emotion speciifc projections
dta_4_plfs_pos_nmf_land_proj_an<- predict(mod_pls_ang,dta_4_plfs_pos_nmf_proj_norm_an , ncomp=17)
dta_4_plfs_pos_nmf_land_proj_sad<- predict(mod_pls_sad,dta_4_plfs_pos_nmf_proj_norm_sad , ncomp=17)
dta_4_plfs_pos_nmf_land_proj_hap<- predict(mod_pls_hap,dta_4_plfs_pos_nmf_proj_norm_hap , ncomp=17)
dta_4_plfs_pos_nmf_proj_norm_an

dta_4_plfs_pos_nmf_land_proj_an<- as.data.frame(dta_4_plfs_pos_nmf_land_proj_an)%>%
  mutate(filename = dta_4_plfs_pos_nmf_proj_norm_an$filename,
         frame =dta_4_plfs_pos_nmf_proj_norm_an$frame,
         expression = dta_4_plfs_pos_nmf_proj_norm_an$expression)

dta_4_plfs_pos_nmf_land_proj_sad<- as.data.frame(dta_4_plfs_pos_nmf_land_proj_sad)%>%
  mutate(filename = dta_4_plfs_pos_nmf_proj_norm_sad$filename,
         frame =dta_4_plfs_pos_nmf_proj_norm_sad$frame,
         expression = dta_4_plfs_pos_nmf_proj_norm_sad$expression)

dta_4_plfs_pos_nmf_land_proj_hap<- as.data.frame(dta_4_plfs_pos_nmf_land_proj_hap)%>%
  mutate(filename = dta_4_plfs_pos_nmf_proj_norm_hap$filename,
         frame =dta_4_plfs_pos_nmf_proj_norm_hap$frame,
         expression = dta_4_plfs_pos_nmf_proj_norm_hap$expression)

dta_4_plfs_pos_nmf_land_proj<- bind_rows(dta_4_plfs_pos_nmf_land_proj_an,
          dta_4_plfs_pos_nmf_land_proj_sad,
          dta_4_plfs_pos_nmf_land_proj_hap)

# spoken emotion specific pls land

dta_4_plfs_spk_nmf_proj_norm

dta_4_plfs_spk_nmf_proj_norm_an<- dta_4_plfs_spk_nmf_proj_norm%>%
  subset(expression == "angry")

dta_4_plfs_spk_nmf_proj_norm_sad<- dta_4_plfs_spk_nmf_proj_norm%>%
  subset(expression == "sad")


dta_4_plfs_spk_nmf_proj_norm_hap<- dta_4_plfs_spk_nmf_proj_norm%>%
  subset(expression == "happy")


# dta_4_plfs_spk_nmf_land_proj<- predict(mod_pls_spoken,dta_4_plfs_spk_nmf_proj_norm , ncomp=17)

dta_4_plfs_spk_nmf_land_proj_an<- predict(mod_pls_spoken_an,dta_4_plfs_spk_nmf_proj_norm_an , ncomp=17)
dta_4_plfs_spk_nmf_land_proj_sad<- predict(mod_pls_spoken_sad,dta_4_plfs_spk_nmf_proj_norm_sad , ncomp=17)
dta_4_plfs_spk_nmf_land_proj_hap<- predict(mod_pls_spoken_hap,dta_4_plfs_spk_nmf_proj_norm_hap , ncomp=17)


dta_4_plfs_spk_nmf_land_proj_an<- as.data.frame(dta_4_plfs_spk_nmf_land_proj_an)%>%
  mutate(filename = dta_4_plfs_spk_nmf_proj_norm_an$filename,
         frame =dta_4_plfs_spk_nmf_proj_norm_an$frame,
         expression = dta_4_plfs_spk_nmf_proj_norm_an$expression)


dta_4_plfs_spk_nmf_land_proj_sad<- as.data.frame(dta_4_plfs_spk_nmf_land_proj_sad)%>%
  mutate(filename = dta_4_plfs_spk_nmf_proj_norm_sad$filename,
         frame =dta_4_plfs_spk_nmf_proj_norm_sad$frame,
         expression = dta_4_plfs_spk_nmf_proj_norm_sad$expression)

dta_4_plfs_spk_nmf_land_proj_hap<- as.data.frame(dta_4_plfs_spk_nmf_land_proj_hap)%>%
  mutate(filename = dta_4_plfs_spk_nmf_proj_norm_hap$filename,
         frame =dta_4_plfs_spk_nmf_proj_norm_hap$frame,
         expression = dta_4_plfs_spk_nmf_proj_norm_hap$expression)

dta_4_plfs_spk_nmf_land_proj<- bind_rows(dta_4_plfs_spk_nmf_land_proj_an,
          dta_4_plfs_spk_nmf_land_proj_sad,
          dta_4_plfs_spk_nmf_land_proj_hap)

```


# now export plfs

continue here

```{r}
dta_4_plfs_spk_nmf_land_proj
dta_4_plfs_pos_nmf_land_proj

# dta_4_plfs_pos_nmf_land_proj<- as.data.frame(dta_4_plfs_pos_nmf_land_proj)

# 
# dta_4_plfs_pos_nmf_land_proj$filename<- dta_4_plfs_pos_nmf_proj$filename
# dta_4_plfs_pos_nmf_land_proj$frame<- dta_4_plfs_pos_nmf_proj$frame


# unique(dta_4_plfs_pos_nmf_land_proj$filename)
dta_4_plfs_pos_nmf_land_proj

dta_4_plfs_pos_nmf_land_proj%>%
  pivot_longer(
    cols = matches("^[xy]_"),
    names_to = c(".value", "id"),
    names_pattern = "([xy])_(\\d+)\\.17 comps"
  ) %>%
  mutate(id = as.integer(id))%>%
  group_by(filename,id,expression)%>%
  summarise_if(is.numeric,mean, na.rm = T)%>%
  ggplot(aes(x,y))+
  geom_point()+
  facet_grid(~expression)




  
# 
# dta_4_plfs_spk_nmf_land_proj<- as.data.frame(dta_4_plfs_spk_nmf_land_proj)
# 
# #
# dta_4_plfs_spk_nmf_land_proj$filename<- dta_4_plfs_spk_nmf_proj$filename
# dta_4_plfs_spk_nmf_land_proj$frame<- dta_4_plfs_spk_nmf_proj$frame
# 
# dta_4_plfs_spk_nmf_land_proj$expression<- dta_4_plfs_spk_nmf_proj$expression


# 

dta_4_plfs_pos_spk_nmf_land_proj<- bind_rows(dta_4_plfs_pos_nmf_land_proj,
          dta_4_plfs_spk_nmf_land_proj
          )


# colnames(dta_4_plfs_pos_spk_nmf_land_proj)


dta_4_plfs_pos_spk_nmf_land_proj<- dta_4_plfs_pos_spk_nmf_land_proj%>%
mutate(time = frame / 25)





```


animate reconstructions
```{r}

tmp_filenames <- unique(dta_4_plfs_pos_spk_nmf_land_proj$filename)

# colnames(dta_4_plfs_pos_spk_nmf_land_proj)


dta_4_plfs_pos_spk_nmf_land_proj_long <- dta_4_plfs_pos_spk_nmf_land_proj %>%
  pivot_longer(
    cols = matches("^[xy]_"),
    names_to = c(".value", "id"),
    names_pattern = "([xy])_(\\d+)\\.17 comps"
  ) %>%
  mutate(id = as.integer(id))  # Convert id to numeric if required

# df_long now has the columns: filename, frame, time, id, x, y
require(gganimate)

for(fn in tmp_filenames) {
  # Filter the data for the current filename.
  d_tmp <- dta_4_plfs_pos_spk_nmf_land_proj_long %>% filter(filename == fn)
    # d_tmp <- dta_4_plfs_pos_spk_nmf_land_proj_long %>% filter(filename == tmp_filenames[1])
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  # no smooth first
  # x_smooth, y = y_smooth
  p <- ggplot(d_tmp, aes(x = x, y = y)) +
    geom_point(color = "white", size = 4) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0(fn, "_nmf.mp4"), animation = anim)
}


```