---
title: "Untitled"
author: "Helio"
date: "2025-03-23"
output: html_document
---

is we wanted to create PLFs I think we can use these below dta_pos_aligned_to_nt_and_eye.rds

dta_eye_align_combined_norm_01_wide_AUS

spoken

tmp_eye_align_data_spk_combined_norm_01_wide dta_spk_eye_align_combined_norm_01_wide_AUS

# Sci Rep

Reviewers want to see some perceptual validation of the stimuli as well as the value of the components (e.g. reconstruction). The latter is beyong the scope bu for the former here is what we can do

select a subset of stimuli (or all)

create PLFs

then present and have participants rate happy, anger, sad, and other (to make it difficult), and neutral can be reconstructed from the lower ratings of these.

Analyse whether the ratings differ depending on the conditions intended

a way to get to the second but is to use correlate model predictions with observers

and also use components to classify participants categorizations derived from above

```{r}
# check the dat that contains landamrks
dta_eye_align_combined_norm_01_wide_AUS
range(dta_eye_align_combined_norm_01_wide_AUS$frame)
range(dta_spk_eye_align_combined_norm_01_wide_AUS$frame)

unique(dta_eye_align_combined_norm_01_wide_AUS$filename)
dta_eye_align_combined_norm_01_wide_AUS

unique(dta_spk_eye_align_combined_norm_01_wide_AUS$filename)

colnames(dta_eye_align_combined_norm_01_wide_AUS)
colnames(dta_spk_eye_align_combined_norm_01_wide_AUS)



```

# combine datasets

```{r}
require(tidyverse)
dta_4_plfs<- bind_rows(dta_eye_align_combined_norm_01_wide_AUS,
          dta_spk_eye_align_combined_norm_01_wide_AUS)


unique(dta_eye_align_combined_norm_01_wide_AUS$filename)

dta_4_plfs<- dta_4_plfs %>%
  # mutate(posed.spoken = if_else(grepl("posed", filename), 
  #                               "posed", "spoken")) %>%
   mutate(parts = str_split(filename, "_", simplify = TRUE)) %>%
  # Create new variables based on the split parts
  mutate(
    posed.spoken = parts[, 2],
    expression   = parts[, 3],
    day          = parts[, 4],
    subject            = parts[, 5]
  )

unique(dta_4_plfs$posed.spoken)
unique(dta_4_plfs$expression)
unique(dta_4_plfs$day)
length(unique(dta_4_plfs$subject))



head(dta_4_plfs)

```

Now create PLFs start win one example

```{r}

setwd("~/Data/scirep_revisions")
require(tidyverse)
require(gganimate)
unique(dta_4_plfs$filename)
# cut_spoken_sad_day2_p31
# cut_posed_angry_day1_p1

colnames(dta_4_plfs)

dta_4_plfs_sel_long
dta_4_plfs_sel_long<-
dta_4_plfs %>%
  select(filename, frame, matches("^(x|y)_\\d+")) %>%
  pivot_longer(
    # id_cols = c(filename, frame),
    cols = matches("^(x|y)_\\d+"),
    names_to = c(".value", "landmark"),
    names_pattern = "([xy])_(\\d+)")


dta_4_plfs_sel_long<- dta_4_plfs_sel_long%>%
mutate(time = frame / 25)
# Assume you have a 'frame' variable in your dataset for the animation timeline.
# If not, you may need to create one.

# Process each unique filename separately.
tmp_filenames <- unique(dta_4_plfs_sel_long$filename)

tmp_filenames

dta_4_plfs_sel_long



dta_4_plfs_sel_long%>%
  group_by(filename)%>%
  mutate(time = frame/25)%>%
  summarise_at(c("frame", "time"), max)%>%
  ggplot(aes(time))+
  geom_histogram()





for(fn in tmp_filenames) {
  # Filter the data for the current filename.
  d_tmp <- dta_4_plfs_sel_long %>% filter(filename == fn)
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  p <- ggplot(d_tmp, aes(x = x, y = y)) +
    geom_point(color = "white", size = 3.5) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0(fn, ".mp4"), animation = anim)
}


# just select test with pratice for 
dta_4_plfs_sel_long$filename
dta_4_plfs_sel_long_test_prct<- dta_4_plfs_sel_long %>% 
  filter(filename %in% c(pos_test_files, spk_test_files,
                         "cut_posed_angry_day1_p1",
                         "cut_spoken_happy_day1_p43"))

nrow(dta_4_plfs_sel_long)
nrow(dta_4_plfs_sel_long_test_prct)

colnames(dta_4_plfs_sel_long_test_prct)
dta_4_plfs_sel_long_test_prct





```

```{r}


# smooth a bit to avoid inherent jitter

#test one jittery video


test1 <- dta_4_plfs_sel_long_test_prct %>%
  group_by(filename, landmark) %>%
  arrange(frame) %>%
  mutate(
    x_smooth = rollmean(x, k = 2, fill = NA, align = "center"),
    y_smooth = rollmean(y, k = 2, fill = NA, align = "center")
  ) %>%
  # Fill NA values at the beginning
  mutate(
    x_smooth = na.locf(x_smooth, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, na.rm = FALSE)
  ) %>%
  # Fill NA values at the end (from last observation backward)
  mutate(
    x_smooth = na.locf(x_smooth, fromLast = TRUE, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, fromLast = TRUE, na.rm = FALSE)
  )%>%
  subset(filename == "cut_spoken_angry_day1_p3")


#setwd("HaloStudy/Data/scirep_revisions/test_vid/smooth")


test1<- test1%>%
mutate(time = frame / 25)

for(fn in unique(test1$filename)) {
  # Filter the data for the current filename.
  d_tmp <- test1 %>% filter(filename == fn)
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  p <- ggplot(d_tmp, aes(x = x_smooth, y = y_smooth)) +
    geom_point(color = "white", size = 3.5) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0("test_smooth_cut_spoken_angry_day1_p3_2f", ".mp4"), animation = anim)
}

# this suggests 2 frames smoothiong is enough
library(dplyr)
library(zoo)

dta_4_plfs_sel_long_test_prct <- dta_4_plfs_sel_long_test_prct %>%
  group_by(filename, landmark) %>%
  arrange(frame) %>%
  mutate(
    x_smooth = rollmean(x, k = 2, fill = NA, align = "center"),
    y_smooth = rollmean(y, k = 2, fill = NA, align = "center")
  ) %>%
  # Fill NA values at the beginning
  mutate(
    x_smooth = na.locf(x_smooth, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, na.rm = FALSE)
  ) %>%
  # Fill NA values at the end (from last observation backward)
  mutate(
    x_smooth = na.locf(x_smooth, fromLast = TRUE, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, fromLast = TRUE, na.rm = FALSE)
  )


#unique(unique(dta_4_plfs_sel_long_test_prct$filename))


tmp_filenames <- unique(dta_4_plfs_sel_long_test_prct$filename)

#setwd("HaloStudy/Data/scirep_revisions/test_vid/smooth")



dta_4_plfs_sel_long_test_prct<- dta_4_plfs_sel_long_test_prct%>%
mutate(time = frame / 25)

for(fn in tmp_filenames) {
  # Filter the data for the current filename.
  d_tmp <- dta_4_plfs_sel_long_test_prct %>% filter(filename == fn)
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  p <- ggplot(d_tmp, aes(x = x_smooth, y = y_smooth)) +
    geom_point(color = "white", size = 3.5) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0(fn, ".mp4"), animation = anim)
}



```

OKAY these are too may stimuli for people to rate

so either we select a random set, or just the test set in classification

saveRDS(pos_test_files, "pos_test_files.rds") saveRDS(spk_test_files,"spk_test_files.rds" )

saveRDS(pos_train_files, "pos_train_files.rds") saveRDS(spk_train_files,"spk_train_files.rds" ) <!--#  -->

```{r}
paths$envRdata
pos_test_files <- readRDS("~/pos_test_files.rds")
spk_test_files <- readRDS("~/spk_test_files.rds")


pos_train_files <- readRDS("~/pos_train_files.rds")
spk_train_files <- readRDS("~/spk_train_files.rds")

pos_test_files
spk_test_files

pos_test_files
dta_4_plfs$filename

pos_test_files<- str_remove(pos_test_files, "./")
pos_test_files<- str_remove(pos_test_files, ".csv")

pos_test_files


spk_test_files<- str_remove(spk_test_files, "./")
spk_test_files<- str_remove(spk_test_files, ".csv")

# train
pos_train_files<- str_remove(pos_train_files, "./")
pos_train_files<- str_remove(pos_train_files, ".csv")

spk_train_files<- str_remove(spk_train_files, "./")
spk_train_files<- str_remove(spk_train_files, ".csv")

spk_test_files

spk_train_files
dta_4_plfs$filename

dta_4_plfs_test_set<- dta_4_plfs %>% 
  filter(filename %in% c(pos_test_files, spk_test_files,
                         "cut_posed_angry_day1_p1",
                         "cut_spoken_happy_day1_p43"))


dta_4_plfs_test_set_agg<-dta_4_plfs_test_set%>%
  group_by(filename, posed.spoken, expression, day, subject)%>%
  summarise_at(c("frame"), max)


dta_4_plfs_test_set_agg$filename_mp4<- paste0(dta_4_plfs_test_set_agg$filename, ".mp4")


setwd("~/scirep_revisions")

dta_4_plfs_test_set_agg$posed.spoken_1_2<- if_else(dta_4_plfs_test_set_agg$posed.spoken == "posed",1,2)

write_csv(dta_4_plfs_test_set_agg, "dta_4_plfs_test_set_agg.csv")


library(readr)
dta_4_plfs_test_set_stim <- read_csv("dta_4_plfs_test_set_stim.csv")

# copy stim to a foler called test_set videos




# Extract the list of video filenames from the dataframe
tmp_vid_list <- unique(dta_4_plfs_test_set_stim$filename_mp4)

na.exclude(tmp_vid_list)

tmp_vid_list<- tmp_vid_list[2:93]

# Define the destination directory
dest_dir <- "test_vid"

# Create the destination directory if it doesn't exist
if (!dir.exists(dest_dir)) {
  dir.create(dest_dir)
}

# Copy the files from the working directory to the destination folder
file.copy(from = tmp_vid_list, to = dest_dir)

```

load behavioural data

```{r}

library(readr)
dta_scirep_grl <- read_csv("data_exp_220644-v4_task-62d6.csv")


library(readr)
dta_scirep_grl_h <- read_csv("data-task-62d6_H.csv")

dta_scirep_grll<- dta_scirep_grl%>%
  janitor::clean_names()%>%
  bind_rows(dta_scirep_grl_h%>%janitor::clean_names())

colnames(dta_scirep_grll)

dta_scirep_grll
dta_scirep_grll1<- dta_scirep_grll%>%
  select(participant_device_type,
         participant_public_id,
         participant_private_id,
          starts_with("spreadsheet"),
         component_name,
         object_name,
         response,
         
         )%>%
subset(grepl("Trial",spreadsheet_display))%>%
subset(grepl("Slider",component_name))%>%
  subset(!is.na(response))%>%
  mutate(response = as.numeric(response))

dta_scirep_grll1%>%
  ggplot(aes(spreadsheet_expression, response, 
             colour  = object_name,
             shape = object_name,
             group = object_name))+
  #geom_point(aes(group = object_name), position = position_dodge(1))+
  facet_grid(~spreadsheet_posed_spoken)+
  stat_summary(geom = "pointrange",position = position_dodge(1))+
  facet_grid(~participant_private_id)


dta_scirep_grll1

dta_scirep_grll1$filename<- dta_scirep_grll1$spreadsheet_filename


#  dta_pos_cfmts_4_pca_test <- readRDS("~//Processeddata/dta_pos_cfmts_4_pca_test.rds")
# > dta_spoken_cfmts_4_pca_test <- readRDS("~Processeddata/dta_spoken_cfmts_4_pca_test.rds")

dta_scirep_grll1$spreadsheet_filename
unique(dta_pos_cfmts_4_pca_test$filename)
unique(dta_spoken_cfmts_4_pca_test$filename)

dta_pos_cfmts_4_pca_test
dta_spoken_cfmts_4_pca_test

dta_rf_test_pos_spk<- bind_rows(dta_pos_cfmts_4_pca_test,
          dta_spoken_cfmts_4_pca_test)

dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,
                                          "./")

dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,".csv")

setdiff(dta_rf_test_pos_spk$filename,dta_scirep_grll1$spreadsheet_filename)

intersect(dta_rf_test_pos_spk$filename,dta_scirep_grll1$spreadsheet_filename)


colnames(dta_rf_test_pos_spk)

dta_rf_test_pos_spk[,c(1:5,120:126)]

dta_scirep_grll1




```

now create categorical ratings

```{r}
dta_scirep_grll1

names(dta_scirep_grll1)<- str_remove(colnames(dta_scirep_grll1), "spreadsheet_")

colnames(dta_scirep_grll1)

# Specify the identifier columns – these are all columns you do not wish to pivot.
id_vars <- c("participant_device_type", "participant_public_id", "participant_private_id", 
             "spreadsheet", "display", "filename", 
             "posed_spoken", "day", "subject", "expression",
             "frame", "filename_mp4", "posed_spoken_1_2", 
             "component_name", "object_name")


dta_scirep_grll1
# Pivot the data: each unique spreadsheet_expression becomes a new column
dta_scirep_grll1$filename
colnames(dta_scirep_grll1)
dta_scirep_grll1<- dta_scirep_grll1[,c(1:16)]

dta_scirep_grll1_wd <- dta_scirep_grll1 %>%
  pivot_wider(
    #id_cols = all_of(id_vars),
    names_from = object_name,
    values_from = response
  )

dta_scirep_grll1_wd

# Create a new column 'max_expression' containing the expression with the highest response

selected_cols<- c("happy_slider", "angry_slider", "sad_slider", "other_emotion")

dta_scirep_grll1_wd <- dta_scirep_grll1_wd %>%
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()

dta_scirep_grll1_wd$max_expression_resp

#aggregated across subjects
colnames(dta_scirep_grll1_wd)
unique(dta_scirep_grll1_wd$component_name)

dta_scirep_grll1_wd_by_vid<-  dta_scirep_grll1_wd%>%
  group_by(display,
  filename, posed_spoken,expression,day,subject,
  filename_mp4, posed_spoken_1_2)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()

#create emotion catgeory by video
colnames(dta_scirep_grll1_wd_by_vid)

dta_scirep_grll1_wd_by_vid <- dta_scirep_grll1_wd_by_vid %>%
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()

dta_scirep_grll1_wd_by_vid$max_expression_resp


dta_rf_test_pos_spk$prediction
dta_scirep_grll1_wd


dta_rf_test_pos_spk$subject
dta_scirep_grll1_wd$subject

dta_rf_test_pos_spk_w_subj<- left_join(dta_scirep_grll1_wd, dta_rf_test_pos_spk,
                                        by = c("filename", "expression"))

dta_scirep_grll1_wd_by_vid$max_expression_resp

dta_rf_test_pos_spk$prediction

dta_scirep_grll1_wd_by_vid$expression
dta_rf_test_pos_spk$expression
dta_scirep_grll1_wd_by_vid$subject
dta_rf_test_pos_spk$subject

dta_rf_test_pos_spk_by_vid<- left_join(dta_scirep_grll1_wd_by_vid, dta_rf_test_pos_spk, 
                                       by = c("filename", "expression"))

sub
# keep just emotion name
dta_rf_test_pos_spk_by_vid$max_expression_resp2<- sub("_.*", "",  dta_rf_test_pos_spk_by_vid$max_expression_resp)


dta_rf_test_pos_spk_w_subj$max_expression_resp2 <- sub("_.*", "",  dta_rf_test_pos_spk_w_subj$max_expression_resp)

# check correspondence
library(caret)
dta_rf_test_pos_spk_by_vid$prediction<- factor(dta_rf_test_pos_spk_by_vid$prediction)

dta_rf_test_pos_spk_by_vid$max_expression_resp2<- factor(dta_rf_test_pos_spk_by_vid$max_expression_resp2)


dta_rf_test_pos_spk_w_subj$max_expression_resp2<- factor(dta_rf_test_pos_spk_w_subj$max_expression_resp2)



dta_rf_test_pos_spk_w_subj$prediction<- factor(dta_rf_test_pos_spk_w_subj$max_expression_resp2)

dta_rf_test_pos_spk_by_vid$prediction
options(scipen = 999)

chisq.test(dta_rf_test_pos_spk_by_vid$prediction,
           dta_rf_test_pos_spk_by_vid$max_expression_resp2)

chisq.test(dta_rf_test_pos_spk_w_subj$prediction,
           dta_rf_test_pos_spk_w_subj$max_expression_resp)


confusionMatrix(dta_rf_test_pos_spk_by_vid$prediction, 
                dta_rf_test_pos_spk_by_vid$max_expression_resp2)

confusionMatrix(dta_rf_test_pos_spk_w_subj$prediction, 
                dta_rf_test_pos_spk_w_subj$max_expression_resp2)


```

Now can we train the model in one participant and test it on another

similarly, can we train it in some data and valite on another within each ppt

lets start with aggregated data and just train with leave one out

```{r}
dta_rf_test_pos_spk_by_vid

?trainControl()
# predict participants ratings from spatiotemp comp data and LOO
mod_rf_spk_mts_train <- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                data = dta_rf_test_pos_spk_by_vid,
                method = "rf",
                trControl = trainControl(method = "oob"),

                ntree = 500) # tunig 

mod_emoperc_nmf<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_rf_test_pos_spk_by_vid,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)

summary(mod_emoperc_nmf)

mod_emoperc_nmf$results

confusionMatrix(data = mod_emoperc_nmf$pred$pred, 
                reference = mod_emoperc_nmf$pred$obs)

unique(dta_rf_test_pos_spk_w_subj$participant_private_id)
# 13017672 (actual ppt)
#13017685 = H

mod_emoperc_nmf2<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_rf_test_pos_spk_w_subj%>%
                           subset(participant_private_id == "13017672"),
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)

summary(mod_emoperc_nmf)

mod_emoperc_nmf2$results

confusionMatrix(data = mod_emoperc_nmf2$pred$pred, 
                reference = mod_emoperc_nmf2$pred$obs)



```

tmp1\<-factor(predict(mod_emoperc_nmf2, dta_rf_test_pos_spk_w_subj%\>% subset(participant_private_id != "13017672")))

tmp2\<-dta_rf_test_pos_spk_w_subj%\>% subset(participant_private_id != "13017672")

tmp2$tmp1<- tmp1
tmp3<- tmp2%>%subset(tmp1!= "other")
confusionMatrix(data = factor(tmp3$tmp1), reference = factor(tmp3\$max_expression_resp2))

```{r}


# Save all specified objects to a file called "project_objects.RData"
save(
  dta_4_plfs,
  dta_4_plfs_sel_long,
  dta_4_plfs_sel_long_test_prct,
  pos_test_files,
  spk_test_files,      # Only include once if there's a duplicate name
  pos_train_files,
  spk_train_files,     # Assuming you meant spk_train_files instead of a repeated spk_test_files
  dta_4_plfs_test_set,
  dta_4_plfs_test_set_agg,
  dta_4_plfs_test_set_stim,
  dta_scirep_grll,
  dta_scirep_grll1,
  dta_rf_test_pos_spk,
  dta_scirep_grll1_wd,
  dta_scirep_grll1_wd_by_vid,
  dta_rf_test_pos_spk_w_subj,
  dta_rf_test_pos_spk_by_vid,
  mod_emoperc_nmf,
  mod_emoperc_nmf2,
  file = "stim_valid_scirep.RData"
)

```
