---
title: "Untitled"
author: "Helio"
date: "2025-03-23"
output: html_document
---

is we wanted to create PLFs I think we can use these below dta_pos_aligned_to_nt_and_eye.rds

dta_eye_align_combined_norm_01_wide_AUS

spoken

tmp_eye_align_data_spk_combined_norm_01_wide dta_spk_eye_align_combined_norm_01_wide_AUS

# Sci Rep

Reviewers want to see some perceptual validation of the stimuli as well as the value of the components (e.g. reconstruction). The latter is beyong the scope bu for the former here is what we can do

select a subset of stimuli (or all)

create PLFs

then present and have participants rate happy, anger, sad, and other (to make it difficult), and neutral can be reconstructed from the lower ratings of these.

Analyse whether the ratings differ depending on the conditions intended

a way to get to the second but is to use correlate model predictions with observers

and also use components to classify participants categorizations derived from above

```{r}
# check the dat that contains landamrks
dta_eye_align_combined_norm_01_wide_AUS
range(dta_eye_align_combined_norm_01_wide_AUS$frame)
range(dta_spk_eye_align_combined_norm_01_wide_AUS$frame)

unique(dta_eye_align_combined_norm_01_wide_AUS$filename)
dta_eye_align_combined_norm_01_wide_AUS

unique(dta_spk_eye_align_combined_norm_01_wide_AUS$filename)

colnames(dta_eye_align_combined_norm_01_wide_AUS)
colnames(dta_spk_eye_align_combined_norm_01_wide_AUS)



```

# combine datasets

```{r}
require(tidyverse)
dta_4_plfs<- bind_rows(dta_eye_align_combined_norm_01_wide_AUS,
          dta_spk_eye_align_combined_norm_01_wide_AUS)


unique(dta_eye_align_combined_norm_01_wide_AUS$filename)

dta_4_plfs<- dta_4_plfs %>%
  # mutate(posed.spoken = if_else(grepl("posed", filename), 
  #                               "posed", "spoken")) %>%
   mutate(parts = str_split(filename, "_", simplify = TRUE)) %>%
  # Create new variables based on the split parts
  mutate(
    posed.spoken = parts[, 2],
    expression   = parts[, 3],
    day          = parts[, 4],
    subject            = parts[, 5]
  )

unique(dta_4_plfs$posed.spoken)
unique(dta_4_plfs$expression)
unique(dta_4_plfs$day)
length(unique(dta_4_plfs$subject))



head(dta_4_plfs)

```

Now create PLFs start win one example

```{r}

setwd("~/Data/scirep_revisions")
require(tidyverse)
require(gganimate)
unique(dta_4_plfs$filename)
# cut_spoken_sad_day2_p31
# cut_posed_angry_day1_p1

colnames(dta_4_plfs)

dta_4_plfs_sel_long
dta_4_plfs_sel_long<-
dta_4_plfs %>%
  select(filename, frame, matches("^(x|y)_\\d+")) %>%
  pivot_longer(
    # id_cols = c(filename, frame),
    cols = matches("^(x|y)_\\d+"),
    names_to = c(".value", "landmark"),
    names_pattern = "([xy])_(\\d+)")


dta_4_plfs_sel_long<- dta_4_plfs_sel_long%>%
mutate(time = frame / 25)
# Assume you have a 'frame' variable in your dataset for the animation timeline.
# If not, you may need to create one.

# Process each unique filename separately.
tmp_filenames <- unique(dta_4_plfs_sel_long$filename)

tmp_filenames

dta_4_plfs_sel_long



dta_4_plfs_sel_long%>%
  group_by(filename)%>%
  mutate(time = frame/25)%>%
  summarise_at(c("frame", "time"), max)%>%
  ggplot(aes(time))+
  geom_histogram()





for(fn in tmp_filenames) {
  # Filter the data for the current filename.
  d_tmp <- dta_4_plfs_sel_long %>% filter(filename == fn)
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  p <- ggplot(d_tmp, aes(x = x, y = y)) +
    geom_point(color = "white", size = 3.5) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0(fn, ".mp4"), animation = anim)
}


# just select test with pratice for 
dta_4_plfs_sel_long$filename
dta_4_plfs_sel_long_test_prct<- dta_4_plfs_sel_long %>% 
  filter(filename %in% c(pos_test_files, spk_test_files,
                         "cut_posed_angry_day1_p1",
                         "cut_spoken_happy_day1_p43"))

nrow(dta_4_plfs_sel_long)
nrow(dta_4_plfs_sel_long_test_prct)

colnames(dta_4_plfs_sel_long_test_prct)
dta_4_plfs_sel_long_test_prct

saveRDS(dta_4_plfs, "dta_4_plfs_orig.Rds")

```

# smooth a bit to avoid inherent jitter

```{r}




#test one jittery video

require(zoo)
test1 <- dta_4_plfs_sel_long_test_prct %>%
  group_by(filename, landmark) %>%
  arrange(frame) %>%
  mutate(
    x_smooth = rollmean(x, k = 2, fill = NA, align = "center"),
    y_smooth = rollmean(y, k = 2, fill = NA, align = "center")
  ) %>%
  # Fill NA values at the beginning
  mutate(
    x_smooth = na.locf(x_smooth, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, na.rm = FALSE)
  ) %>%
  # Fill NA values at the end (from last observation backward)
  mutate(
    x_smooth = na.locf(x_smooth, fromLast = TRUE, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, fromLast = TRUE, na.rm = FALSE)
  )%>%
  subset(filename == "cut_spoken_happy_day1_p43")


#setwd("HaloStudy/Data/scirep_revisions/test_vid/smooth")




  
  # Calculate duration for this file from the time variable.
  # file_duration <- max(d_tmp$time)
  # cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
# tmp_plf <- ggplot(d_tmp, aes(x = x_smooth, y = y_smooth)) +
#     geom_point(color = "white", size = 3.5) +
#     scale_y_reverse()+
#     coord_fixed() +
#     theme_void() +
#     theme(
#       plot.background = element_rect(fill = "black", colour = NA),
#       panel.background = element_rect(fill = "black", colour = NA),
#       plot.margin = margin(0, 0, 0, 0)
#     ) +
#     transition_time(time) +
#     labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
#   
#   # Render the animation as an MP4 with 24 fps.
#   anim <- animate(p, 
#                   width = 1920, 
#                   height = 1080, 
#                   # fps = 24,
#                   duration = file_duration,  # per-file duration
#                   renderer = ffmpeg_renderer(),
#                   bg = "black")
#   
#   # Save the animation.
#   anim_save(paste0("test_smooth_cut_spoken_angry_day1_p3_2f", ".mp4"), animation = anim)
# }



# export fpor presentation


require(ggforce)
  # Filter the data for the current filename.
  d_tmp <- test1 

  test1<- test1%>%
mutate(time = frame / 24)

  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # C
  # require(ggvoronoi)
tmp_plot <-ggplot(d_tmp, aes(x = x_smooth, y = y_smooth)) +
    geom_point(color = "orange", size = 4) +
    scale_y_reverse()+
  # ggforce::geom_delaunay_segment(colour = "gray", alpha =.1)+
    # coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA)
      # plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(tmp_plot, 
                  # width = 1920, 
                  # height = 1080, 
                  fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0("demo_plf_rec", ".mp4"), animation = anim)



```



```{r}

# this suggests 2 frames smoothiong is enough
library(dplyr)
library(zoo)

dta_4_plfs_sel_long_test_prct <- dta_4_plfs_sel_long_test_prct %>%
  group_by(filename, landmark) %>%
  arrange(frame) %>%
  mutate(
    x_smooth = rollmean(x, k = 2, fill = NA, align = "center"),
    y_smooth = rollmean(y, k = 2, fill = NA, align = "center")
  ) %>%
  # Fill NA values at the beginning
  mutate(
    x_smooth = na.locf(x_smooth, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, na.rm = FALSE)
  ) %>%
  # Fill NA values at the end (from last observation backward)
  mutate(
    x_smooth = na.locf(x_smooth, fromLast = TRUE, na.rm = FALSE),
    y_smooth = na.locf(y_smooth, fromLast = TRUE, na.rm = FALSE)
  )


#unique(unique(dta_4_plfs_sel_long_test_prct$filename))


tmp_filenames <- unique(dta_4_plfs_sel_long_test_prct$filename)

#setwd("HaloStudy/Data/scirep_revisions/test_vid/smooth")



dta_4_plfs_sel_long_test_prct<- dta_4_plfs_sel_long_test_prct%>%
mutate(time = frame / 25)

for(fn in tmp_filenames) {
  # Filter the data for the current filename.
  d_tmp <- dta_4_plfs_sel_long_test_prct %>% filter(filename == fn)
  
  # Calculate duration for this file from the time variable.
  file_duration <- max(d_tmp$time)
  cat("Filename:", fn, "- Duration:", file_duration, "\n")
  
  # Create the animation plot.
  p <- ggplot(d_tmp, aes(x = x_smooth, y = y_smooth)) +
    geom_point(color = "white", size = 4) +
    scale_y_reverse()+
    coord_fixed() +
    theme_void() +
    theme(
      plot.background = element_rect(fill = "black", colour = NA),
      panel.background = element_rect(fill = "black", colour = NA),
      plot.margin = margin(0, 0, 0, 0)
    ) +
    transition_time(time) +
    labs(title = paste("Filename:", fn, "- Time: {frame_time}s"))
  
  # Render the animation as an MP4 with 24 fps.
  anim <- animate(p, 
                  width = 1920, 
                  height = 1080, 
                  # fps = 24,
                  duration = file_duration,  # per-file duration
                  renderer = ffmpeg_renderer(),
                  bg = "black")
  
  # Save the animation.
  anim_save(paste0(fn, ".mp4"), animation = anim)
}



```

OKAY these are too may stimuli for people to rate

so either we select a random set, or just the test set in classification

saveRDS(pos_test_files, "pos_test_files.rds") saveRDS(spk_test_files,"spk_test_files.rds" )

saveRDS(pos_train_files, "pos_train_files.rds") saveRDS(spk_train_files,"spk_train_files.rds" ) <!--#  -->

```{r}
paths$envRdata
pos_test_files <- readRDS("~/pos_test_files.rds")
spk_test_files <- readRDS("~/spk_test_files.rds")


pos_train_files <- readRDS("~/pos_train_files.rds")
spk_train_files <- readRDS("~/spk_train_files.rds")

pos_test_files
spk_test_files

pos_test_files
dta_4_plfs$filename

pos_test_files<- str_remove(pos_test_files, "./")
pos_test_files<- str_remove(pos_test_files, ".csv")

pos_test_files


spk_test_files<- str_remove(spk_test_files, "./")
spk_test_files<- str_remove(spk_test_files, ".csv")

# train
pos_train_files<- str_remove(pos_train_files, "./")
pos_train_files<- str_remove(pos_train_files, ".csv")

spk_train_files<- str_remove(spk_train_files, "./")
spk_train_files<- str_remove(spk_train_files, ".csv")

spk_test_files

spk_train_files
dta_4_plfs$filename

dta_4_plfs_test_set<- dta_4_plfs %>% 
  filter(filename %in% c(pos_test_files, spk_test_files,
                         "cut_posed_angry_day1_p1",
                         "cut_spoken_happy_day1_p43"))


dta_4_plfs_test_set_agg<-dta_4_plfs_test_set%>%
  group_by(filename, posed.spoken, expression, day, subject)%>%
  summarise_at(c("frame"), max)


dta_4_plfs_test_set_agg$filename_mp4<- paste0(dta_4_plfs_test_set_agg$filename, ".mp4")


setwd("~/scirep_revisions")

dta_4_plfs_test_set_agg$posed.spoken_1_2<- if_else(dta_4_plfs_test_set_agg$posed.spoken == "posed",1,2)

write_csv(dta_4_plfs_test_set_agg, "dta_4_plfs_test_set_agg.csv")


library(readr)
dta_4_plfs_test_set_stim <- read_csv("dta_4_plfs_test_set_stim.csv")

# copy stim to a foler called test_set videos




# Extract the list of video filenames from the dataframe
tmp_vid_list <- unique(dta_4_plfs_test_set_stim$filename_mp4)

na.exclude(tmp_vid_list)

tmp_vid_list<- tmp_vid_list[2:93]

# Define the destination directory
dest_dir <- "test_vid"

# Create the destination directory if it doesn't exist
if (!dir.exists(dest_dir)) {
  dir.create(dest_dir)
}

# Copy the files from the working directory to the destination folder
file.copy(from = tmp_vid_list, to = dest_dir)

```

# Load behavioral data

this is data from participants rating the test set for emotion

# rejections

No video face detected in front of the computer

63e6378e04ade071ebf9ff9c

```{r}

library(readr)
dta_scirep_grl <- read_csv("data_exp_220644-v4_task-62d6.csv")

library(readr)
# dta_scirep_grl_h <- read_csv("data-task-62d6_H.csv")
dta_scirep_grll<- dta_scirep_grl%>%
  janitor::clean_names()%>%
  bind_rows(dta_scirep_grl_h%>%janitor::clean_names())


unqiue(dta_scirep_grl_s1$participant_private_id)
dta_scirep_grl_s1<-dta_scirep_grll

# prolific data
library(readr)
dta_scirep_all <- read_csv("data_exp_220644-v6_task-62d6.csv")
View(dta_scirep_all)



dta_scirep_all<- dta_scirep_all%>%
  janitor::clean_names()%>%
  bind_rows(dta_scirep_grll%>%janitor::clean_names())
dta_scirep_grll


unique(dta_scirep_all$participant_private_id)

dta_scirep_all<-dta_scirep_all%>%
  subset(participant_private_id!= "13017685")%>%
  subset(!is.na(participant_private_id))


# select necessary columns and rows (e.g. trial, no practice)

unique(dta_scirep_all$task_version)
unique(dta_scirep_all$participant_private_id)


dta_scirep_all2<- dta_scirep_all%>%
  select(
    #participant_device_type,
         participant_public_id,
         participant_private_id,
          starts_with("spreadsheet"),
         trial_number,
         component_name,
         object_name,
         response)%>%
subset(grepl("Trial",spreadsheet_display))%>%
subset(grepl("Slider",component_name))%>%
  subset(!is.na(response))%>%
  mutate(response = as.numeric(response))

colnames(dta_scirep_all2)

names(dta_scirep_all2)<- str_remove(colnames(dta_scirep_all2), "spreadsheet_")
colnames(dta_scirep_all2)

# quick plots of emotion rating
dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  #subset(object_name!= "other_emotion")%>%
  group_by(expression, posed_spoken)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = object_name,
             shape = object_name))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  geom_point(aes(group = object_name), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1))+
  facet_grid(~posed_spoken)

# quick plots of emotion rating
dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  #mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  #subset(object_name!= "other_emotion")%>%
  group_by(expression, posed_spoken)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = object_name,
             shape = object_name))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  #geom_point(aes(group = object_name), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1))+
  facet_grid(~posed_spoken)





dta_scirep_all2$slider <- str_remove(dta_scirep_all2$object_name, "_slider")

dta_scirep_all2$condition <- if_else(dta_scirep_all2$posed_spoken == "posed", "expression only", "emotive speech")

dta_scirep_all2$condition<- factor(dta_scirep_all2$condition, levels = c("expression only", "emotive speech"))

dta_scirep_all2$filename
reviews$stiom_valid<-

dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  # group_by(participant_private_id,expression, object_name, condition,slider)%>%
  group_by(filename,expression, object_name, condition,slider)%>%  
  
  summarise_if(is.numeric, mean, na.rm = T)%>%
  #mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  subset(object_name!= "other_emotion")%>%
  group_by(expression, condition)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = slider,
             shape = slider))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  geom_point(aes(group = slider), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1), size = .6)+
  facet_wrap(~condition, scale = "free_y")+
  scale_color_brewer(palette = "Dark2")+
  ylab("response (z)")+
  p$graphstyle_int+
  theme(axis.text.x = element_text(hjust = 1, angle =45),
        legend.position = "top")


reviews$stiom_valid

ggsave("sm_stim_valid.tiff",
       device = 'tiff', reviews$stiom_valid,
       width = 10,
       height = 6,
       dpi = 300,
       compression = "lzw")



```



# plot distributions

outlier tests etc

```{r}


library(dplyr)
library(ggplot2)
library(tidyr)


dta_scirep_all

dta_scirep_all2 %>%
  filter(participant_public_id != "63e6378e04ade071ebf9ff9c") %>%
  group_by(participant_private_id, expression, object_name, posed_spoken) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  mutate(response = if_else(object_name == "other_emotion" & 
                              response > 75, NA_real_, response)) %>%
  #filter(object_name != "other_emotion") %>%
  group_by(expression, posed_spoken) %>%
  mutate(response = scale(response)) %>%
  ggplot(aes(x = expression, y = response, colour = object_name, shape = object_name)) +
    geom_violin(trim=FALSE,position = position_dodge(1))+
  #geom_boxplot(width = 0.1, position = position_dodge(1)) +
  stat_summary(aes(group = object_name), geom = "pointrange", position = position_dodge(1)) +
  facet_wrap(~posed_spoken, nrow = 2) +
  theme_minimal()


colnames(dta_scirep_all2)
unique(dta_scirep_all2$is_outlier)


dta_scirep_all2 <- dta_scirep_all2 %>%
  filter(!is.na(response)) %>%
  group_by(expression, posed_spoken) %>%
  mutate(
    q_1 = quantile(response, 0.25, na.rm = TRUE),
    q_3 = quantile(response, 0.75, na.rm = TRUE),
    iqr =  q_3-q_1,
    lower_bound = q_1 - 1 * iqr,
    upper_bound = q_3 + 1 * iqr,
    is_outlier = response < lower_bound | response > upper_bound
  ) %>%
  ungroup()

unique(dta_scirep_all2$is_outlier)

table(dta_scirep_all2$is_outlier)

dta_scirep_all2$is_outlier2<-NULL
dta_scirep_all2
dta_scirep_all2%>%
  subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  subset(participant_private_id !="13017685")%>%
     # subset(is_outlier == FALSE)%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%

  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
  subset(posed_spoken == "posed")%>%
  
subset(object_name != "other_emotion")%>%
  ggplot(aes(object_name, response, 
             colour  = object_name))+
    stat_summary(geom = "pointrange",
                 position = position_dodge(1))+
  facet_wrap(~expression, scales = "free_y")



dta_scirep_all2%>%
  subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  subset(participant_private_id !="13017685")%>%
     # subset(is_outlier == FALSE)%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%

  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
  subset(posed_spoken != "posed")%>%
  
subset(object_name != "other_emotion")%>%
  ggplot(aes(object_name, response, 
             colour  = object_name))+
    stat_summary(geom = "pointrange",
                 position = position_dodge(1))+
  facet_wrap(~expression, scales = "free_y")



```

load the test rf data

```{r}
#  dta_pos_cfmts_4_pca_test <- readRDS("~//Processeddata/dta_pos_cfmts_4_pca_test.rds")
# > dta_spoken_cfmts_4_pca_test <- readRDS("~Processeddata/dta_spoken_cfmts_4_pca_test.rds")

unique(dta_pos_cfmts_4_pca_test$filename)
unique(dta_spoken_cfmts_4_pca_test$filename)

dta_pos_cfmts_4_pca_test
dta_spoken_cfmts_4_pca_test

# bind expression only and emotive speech
dta_rf_test_pos_spk<- bind_rows(dta_pos_cfmts_4_pca_test,
          dta_spoken_cfmts_4_pca_test)

# remove patterns to match the filename pattern in the behavioural data
dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,
                                          "./")

dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,".csv")

# check that we have the same filenames in rf and behavioural data
setdiff(dta_rf_test_pos_spk$filename,dta_scirep_all2$filename)
intersect(dta_rf_test_pos_spk$filename,dta_scirep_all2$filename)



```

# now create categorical ratings for emotion

We will assign a category based on the maximum rating, we can ignore neutral in here since we are focusing on test set, and we already use other to make the rating difficult

```{r}

colnames(dta_scirep_all2)

dta_scirep_all2<- dta_scirep_all2[,c(1:16)]

# Specify the identifier columns – these are all columns you do not wish to pivot.
id_vars <- c("participant_device_type", "participant_public_id", "participant_private_id", 
             "spreadsheet", "display", "filename", 
             "posed_spoken", "day", "subject", "expression",
             "frame", "filename_mp4", "posed_spoken_1_2", 
             "component_name", "object_name")


# Pivot the data: each unique spreadsheet_expression becomes a new column

dta_scirep_all_wd <- dta_scirep_all2 %>%
  pivot_wider(
    #id_cols = all_of(id_vars),
    names_from = object_name,
    values_from = response
  )

dta_scirep_all_wd

# Create a new column 'max_expression' containing the expression with the highest response

selected_cols<- c("happy_slider", "angry_slider", "sad_slider", "other_emotion")

dta_scirep_all_wd <- dta_scirep_all_wd %>%
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()%>%
    rowwise() %>%
  mutate(max_expression_resp_n_oth = {
    values <- c_across(all_of(selected_cols[1:3]))
    selected_cols[1:3][which.max(values)]
  }) %>%
    ungroup()

unique(dta_scirep_all_wd$max_expression_resp_n_oth)
table(dta_scirep_all_wd$max_expression_resp)


#aggregated across subjects


dta_scirep_all_wd_by_vid<-  dta_scirep_all_wd%>%
  group_by(display,
  filename, posed_spoken,expression,day,subject,
  filename_mp4, posed_spoken_1_2)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
#create emotion catgeory by video
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()%>%
  rowwise() %>%
  mutate(max_expression_resp_n_oth = {
    values <- c_across(all_of(selected_cols[1:3]))
    selected_cols[1:3][which.max(values)]
  }) %>%
    ungroup()

dta_scirep_all_wd_by_vid$max_expression_resp
dta_scirep_all_wd_by_vid$max_expression_resp_n_oth

# now merge the nnf classification data with subject level perceptual data on the test set
dta_scirep_all_wd
dta_scirep_all_wd_by_sbj<- left_join(dta_scirep_all_wd, dta_rf_test_pos_spk,
                                        by = c("filename", "expression"))



# now merge the nmf classification data with subject aggregated  perceptual data on the test set

dta_scirep_all_wd_by_vid


dta_scirep_all_wd_by_vid<- left_join(dta_scirep_all_wd_by_vid, dta_rf_test_pos_spk, 
                                       by = c("filename", "expression"))

# keep just emotion name
dta_scirep_all_wd_by_vid$max_expression_resp2<- sub("_.*", "",  dta_scirep_all_wd_by_vid$max_expression_resp)


dta_scirep_all_wd_by_sbj$max_expression_resp2 <- sub("_.*", "",  dta_scirep_all_wd_by_sbj$max_expression_resp)


dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2<- sub("_.*", "",  dta_scirep_all_wd_by_vid$max_expression_resp_n_oth)


dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2 <- sub("_.*", "",  dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth)






dta_scirep_all_wd

write_csv(dta_scirep_all_wd, "dta_scirep_all_wd.csv")



dta_scirep_all_wd$max_expression_resp_n_oth




```

Now check correspondence of defined emotion categories and perceptual rating derived categories

```{r}
library(caret)
# check correspondence

options(scipen = 999)

# check factors first
dta_scirep_all_wd_by_vid$prediction<- factor(dta_scirep_all_wd_by_vid$prediction)

dta_scirep_all_wd_by_vid$max_expression_resp2<- factor(dta_scirep_all_wd_by_vid$max_expression_resp2)


dta_scirep_all_wd_by_sbj$max_expression_resp2<- factor(dta_scirep_all_wd_by_sbj$max_expression_resp2)


dta_scirep_all_wd_by_sbj$prediction<- factor(dta_scirep_all_wd_by_sbj$prediction)



```

tabs

```{r}


#simple chi-square test
chisq.test(dta_scirep_all_wd_by_sbj$expression,
           dta_scirep_all_wd_by_sbj$max_expression_resp2)
chisq.test(dta_scirep_all_wd_by_sbj$expression,
           dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2)

#consistent
chisq.test(dta_scirep_all_wd_by_vid$expression,
           dta_scirep_all_wd_by_vid$max_expression_resp2)

chisq.test(dta_scirep_all_wd_by_vid$expression,
           dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2)



```

# confusion matrices between prediction from NMF model and perceptual categorisations

```{r}


library(caret)
# confusion matrix

unique(dta_scirep_all_wd_by_sbj$prediction)
unique(dta_scirep_all_wd_by_sbj$max_expression_resp2)


# 
# unique(dta_scirep_all_wd_by_sbj$prediction)
# unique(dta_scirep_all_wd_by_sbj$max_expression_resp2)

dta_scirep_all_wd_by_vid$prediction<- factor(dta_scirep_all_wd_by_vid$prediction)

dta_scirep_all_wd_by_vid$max_expression_resp2<- factor(dta_scirep_all_wd_by_vid$max_expression_resp2)

dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2<- factor(dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2)
dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2<-factor(dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2)

dta_scirep_all_wd_by_sbj$expression<- factor(dta_scirep_all_wd_by_sbj$expression)


confusionMatrix(dta_scirep_all_wd_by_sbj$expression, 
                dta_scirep_all_wd_by_sbj$max_expression_resp2)

confusionMatrix(dta_scirep_all_wd_by_sbj$expression, 
                dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2)

dta_scirep_all_wd_by_vid$expression<- factor(dta_scirep_all_wd_by_vid$expression)


confusionMatrix(dta_scirep_all_wd_by_vid$expression, 
                dta_scirep_all_wd_by_vid$max_expression_resp2)

confusionMatrix(dta_scirep_all_wd_by_vid$prediction, 
                dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2)


dta_scirep_all_wd_by_vid$prediction


dta_scirep_all_wd_by_vid

dta_scirep_all_wd_by_vid$filename
dta_scirep_all_wd_by_vid$participant_private_id

reviews$sm_nmf_vs_emo_rat <-
dta_scirep_all_wd_by_sbj%>%
  select(filename, participant_private_id,posed_spoken,
         expression,prediction, contains("slider"),max_expression_resp_n_oth2)%>%
pivot_longer(
    cols = c(happy_slider, angry_slider, sad_slider),
    names_to = "slider",
    values_to = "response"
  )%>%
  mutate(condition = posed_spoken)%>%
  mutate(slider = str_remove(slider, "_slider"))%>%

  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  # group_by(participant_private_id,expression, object_name, condition,slider)%>%
  group_by( filename, prediction, condition,slider,max_expression_resp_n_oth2)%>%  
  
  summarise_if(is.numeric, mean, na.rm = T)%>%
  #mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  # subset(object_name!= "other_emotion")%>%
  group_by(prediction, condition)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(prediction, response, 
             colour  = slider,
             shape = slider))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  geom_point(aes(group = slider), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1), size = .6)+
  facet_wrap(~condition, scale = "free_y")+
  scale_color_brewer(palette = "Dark2")+
  xlab("expression prediction from NMF model")+
  ylab("response (z)")+
  p$graphstyle_int+
  theme(axis.text.x = element_text(hjust = 1, angle =45),
        legend.position = "top")


ggsave(reviews$sm_nmf_vs_emo_rat)

ggsave("sm_sm_nmf_vs_emo_rat.tiff",
       device = 'tiff', reviews$sm_nmf_vs_emo_rat,
       width = 10,
       height = 6,
       dpi = 300,
       compression = "lzw")

```

split by posed and spoken

```{r}
#split by posed and spoken
dta_scirep_all_wd_by_sbj_pos<- dta_scirep_all_wd_by_sbj%>%subset(posed_spoken == "posed")
dta_scirep_all_wd_by_sbj_spk<- dta_scirep_all_wd_by_sbj%>%subset(posed_spoken != "posed")

dta_scirep_all_wd_by_vid_pos<- dta_scirep_all_wd_by_vid%>%subset(posed_spoken == "posed")
dta_scirep_all_wd_by_vid_spk<- dta_scirep_all_wd_by_vid%>%subset(posed_spoken != "posed")


# by subj

dta_scirep_all_wd_by_sbj_pos$expression<- factor(dta_scirep_all_wd_by_sbj_pos$expression)



chisq.test(dta_scirep_all_wd_by_sbj_pos$expression, 
                dta_scirep_all_wd_by_sbj_pos$max_expression_resp_n_oth2,
           simulate.p.value = TRUE)

confusionMatrix(dta_scirep_all_wd_by_sbj_pos$expression, 
                dta_scirep_all_wd_by_sbj_pos$max_expression_resp_n_oth2)


confusionMatrix(dta_scirep_all_wd_by_sbj_spk$expression, 
                dta_scirep_all_wd_by_sbj_spk$max_expression_resp_n_oth2)

chisq.test(dta_scirep_all_wd_by_sbj_spk$expression, 
                dta_scirep_all_wd_by_sbj_spk$max_expression_resp_n_oth2,
           simulate.p.value = TRUE)




dta_scirep_all2

```


Use video as we are validating video

```{r}
# by vid
# confusionMatrix(dta_scirep_all_wd_by_vid_pos$prediction, 
#                 dta_scirep_all_wd_by_vid_pos$max_expression_resp2)
dta_scirep_all_wd_by_vid_pos
dta_scirep_all_wd_by_vid_pos$expression<- factor(dta_scirep_all_wd_by_vid_pos$expression)

chisq.test(dta_scirep_all_wd_by_vid_pos$expression, 
                dta_scirep_all_wd_by_vid_pos$max_expression_resp_n_oth2,
           simulate.p.value = TRUE)

chisq.test(dta_scirep_all_wd_by_vid_spk$expression, 
                dta_scirep_all_wd_by_vid_spk$max_expression_resp_n_oth2, simulate.p.value = TRUE)


confusionMatrix(dta_scirep_all_wd_by_vid_pos$expression, 
                dta_scirep_all_wd_by_vid_pos$max_expression_resp_n_oth2)



confusionMatrix(dta_scirep_all_wd_by_vid_spk$expression, 
                dta_scirep_all_wd_by_vid_spk$max_expression_resp_n_oth2)

# dta_scirep_all_wd_by_vid_spk$expression<- factor(dta_scirep_all_wd_by_vid_spk$expression)


```


# report
```{r}
dta_scirep_all_wd_by_vid_spk$expression<- factor(dta_scirep_all_wd_by_vid_spk$expression)
dta_scirep_all_wd_by_vid_spk$max_expression_resp_n_oth2<- factor(dta_scirep_all_wd_by_vid_spk$max_expression_resp_n_oth2)



chisq.test(dta_scirep_all_wd_by_vid_pos$prediction,
           dta_scirep_all_wd_by_vid_pos$max_expression_resp_n_oth2,simulate.p.value = TRUE)

chisq.test(dta_scirep_all_wd_by_vid_spk$prediction,
           dta_scirep_all_wd_by_vid_spk$max_expression_resp_n_oth2,simulate.p.value = TRUE)




```

# Now can we train a model do categorise participants emotion ratings from NMF data

-   do this on aggregated data first (loo)

-   then do it on subject based leaving one participant at a time

similarly, can we train it in some data and valite on another within each ppt

lets start with aggregated data and just train with leave one out

```{r}
dta_scirep_all_wd_by_vid

# predict participants ratings from spatiotemp comp data and LOO
?caret::train()

dta_scirep_all_wd_by_vid$posed_spoken_1_2

mod_emoperc_nmf<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)


mod_emoperc_nmf$results

# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf$pred$pred, 
                reference = mod_emoperc_nmf$pred$obs)


mod_emoperc_nmf_no_oth<- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)


mod_emoperc_nmf_no_oth$results



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_no_oth$pred$pred, 
                reference = mod_emoperc_nmf_no_oth$pred$obs)
#sign






```

By video
SPLIT BY EMOTION ONLY AND EMOTIVE SPEECH

```{r eval=FALSE, include=FALSE}


#pose vs spk

mod_emoperc_nmf_pos<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid_pos,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_pos$pred$pred, 
                reference = mod_emoperc_nmf_pos$pred$obs)


mod_emoperc_nmf_spk <- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid_spk,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_spk$pred$pred, 
                reference = mod_emoperc_nmf_spk$pred$obs)


```

no other
```{r}

mod_emoperc_nmf_pos_no_oth<- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid_pos,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
mod_emoperc_nmf_pos_no_oth$results

confusionMatrix(data = mod_emoperc_nmf_pos_no_oth$pred$pred, 
                reference = mod_emoperc_nmf_pos_no_oth$pred$obs)



```


#spk

```{r}
ctrl_pos_vid <- trainControl(method = "cv", index = loo_train_indices_spk,
                       savePredictions = "final")

set.seed(42)
mod_emoperc_nmf_spk_no_oth <- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_vid_spk,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV",savePredictions = "final"),
                              
                              ntree = 500)


mod_emoperc_nmf_spk_no_oth$pred
# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_spk_no_oth$pred$pred, 
                reference = mod_emoperc_nmf_spk_no_oth$pred$obs)



```



```{r}


unique(dta_rf_test_pos_spk_w_subj$participant_private_id)
# 13017672 (actual ppt)
#13017685 = H



mod_emoperc_nmf_spk$finalModel$forest

  mod_emoperc_nmf_spk$pred %>%
  group_by(rowIndex) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(rowIndex), accuracy))+
  geom_point()+
  geom_hline(yintercept = 0.4554)+
  geom_bar(stat = "identity", width = .1)+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()
  
  
  

  
  
  mod_emoperc_nmf_pos_no_oth %>%
  group_by(rowIndex) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(rowIndex), accuracy))+
  geom_point()+
    
    geom_hline(yintercept = 0.4554, colour = "red", linetype = "dashed")+
 geom_hline(yintercept = .33, colour = "pink", linetype = "dashed")+
    
    
  geom_bar(stat = "identity", width = .1)+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()
  

# dta_scirep_all_wd_by_vid_spk$prediction_nmf_comp_rat<- predict(mod_emoperc_nmf_spk_no_oth,dta_scirep_all_wd_by_vid_spk)


# dta_scirep_all_wd_by_vid_pos$prediction_nmf_comp_rat<- predict(mod_emoperc_nmf_pos_no_oth,dta_scirep_all_wd_by_vid_pos)


# these give exactly the same result
  
  
  
  predictions_df <- mod_emoperc_nmf_spk_no_oth$pred

# Identify whether each prediction was correct
predictions_df$correct <- ifelse(predictions_df$pred == predictions_df$obs, 1, 0)

# Create an index to represent each left-out sample (fold)
predictions_df$fold_index <- seq_len(nrow(predictions_df))

# Plot using ggplot2
library(ggplot2)

ggplot(predictions_df, aes(x = fold_index, y = correct)) +
  geom_point() +
  geom_line() +
  ylim(0, 1) +
  labs(x = "Left-out Sample (Fold Index)", 
       y = "Accuracy (0 = Incorrect, 1 = Correct)",
       title = "LOOCV Accuracy for Each Left-out Sample")


```

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)

dta_scirep_all_wd_by_vid_spk


# Updated helper function to generate decision boundary plots with overlaid data points
plot_decision_boundary <- function(data, model, comp_x, comp_y, fixed_comp, fixed_value, 
                                   x_label = comp_x, y_label = comp_y) {
  # Create a sequence over the range of the two components
  x_range <- seq(min(data[[comp_x]], na.rm = TRUE), max(data[[comp_x]], na.rm = TRUE), length.out = 100)
  y_range <- seq(min(data[[comp_y]], na.rm = TRUE), max(data[[comp_y]], na.rm = TRUE), length.out = 100)
  
  # Create grid for predictions using base R naming
  grid <- expand.grid(x_range, y_range)
  colnames(grid) <- c(comp_x, comp_y)
  grid[[fixed_comp]] <- fixed_value
  
  # Predict using the model
  grid$pred <- predict(model, newdata = grid)
  
  # Generate the plot
  p <- ggplot() +
    # Decision boundaries from prediction grid
    geom_tile(data = grid, aes_string(x = comp_x, y = comp_y, fill = "pred"), alpha = 0.2) +
    # stat_contour(data = grid, aes_string(x = comp_x, y = comp_y, z = "as.numeric(pred)"),
    #              color = "black", bins = 3, alpha = 0.7) +
    # Overlay actual data points with true labels
    geom_point(data = data, 
               aes_string(x = comp_x, y = comp_y, colour = "expression"), 
               size = 2) +
    labs(title = paste(x_label, "vs.", y_label),
         x = x_label,
         y = y_label) +
    theme_minimal()
  
  return(p)
}

# Use the Emotive Speech data and the trained model (mod_emoperc_nmf_spk_no_oth)
data_spk <- dta_scirep_all_wd_by_vid_spk

# For comp1 vs comp2, hold comp3 constant at its median value:
p1 

test<-data_spk%>%
                         subset(comp1_scale>-3 & comp2_scale<3)

plot_decision_boundary(data = test,
                       model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp1_scale", comp_y = "comp2_scale", 
                             fixed_comp = "comp3_scale", 
                             fixed_value = median(test$comp3_scale, na.rm = TRUE),
                             x_label = "Component 1", y_label = "Component 2")

# For comp1 vs comp3, hold comp2 constant at its median value:
plot_decision_boundary(data = test, model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp1_scale", comp_y = "comp3_scale", 
                             fixed_comp = "comp2_scale", 
                             fixed_value = median(test$comp2_scale, na.rm = TRUE),
                             x_label = "Component 1", y_label = "Component 3")

# For comp2 vs comp3, hold comp1 constant at its median value:
 plot_decision_boundary(data = test, model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp2_scale", comp_y = "comp3_scale", 
                             fixed_comp = "comp1_scale", 
                             fixed_value = median(test$comp1_scale, na.rm = TRUE),
                             x_label = "Component 2", y_label = "Component 3")

# Arrange the plots in a grid
grid.arrange(p1, p2, p3, ncol = 2)


plot_decision_boundary(data = dta_scirep_all_wd_by_vid_pos,
                       model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp1_scale", comp_y = "comp2_scale", 
                             fixed_comp = "comp3_scale", 
                             fixed_value = median(dta_scirep_all_wd_by_vid_pos$comp3_scale, na.rm = TRUE),
                             x_label = "Component 1", y_label = "Component 2")

# For comp1 vs comp3, hold comp2 constant at its median value:
plot_decision_boundary(data = dta_scirep_all_wd_by_vid_pos, model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp1_scale", comp_y = "comp3_scale", 
                             fixed_comp = "comp2_scale", 
                             fixed_value = median(dta_scirep_all_wd_by_vid_pos$comp2_scale, na.rm = TRUE),
                             x_label = "Component 1", y_label = "Component 3")

# For comp2 vs comp3, hold comp1 constant at its median value:
 plot_decision_boundary(data = dta_scirep_all_wd_by_vid_pos, model = mod_emoperc_nmf_spk_no_oth,
                             comp_x = "comp2_scale", comp_y = "comp3_scale", 
                             fixed_comp = "comp1_scale", 
                             fixed_value = median(dta_scirep_all_wd_by_vid_pos$comp1_scale, na.rm = TRUE),
                             x_label = "Component 2", y_label = "Component 3")

                          
                          
 

```

by subject

where leave one out is the subject (to avoid data leakage

```{r}


library(caret)
library(dplyr)

dta_scirep_all_wd_by_sbj_pos
participants <- unique(dta_scirep_all_wd_by_sbj$participant_private_id)

# Create a list of training indices for leave-one-participant-out:
# For each participant, the training set consists of all rows not belonging to that participant.
loo_train_indices <- lapply(participants, function(p) {
  which(dta_scirep_all_wd_by_sbj$participant_private_id != p)
})
names(loo_train_indices) <- participants

# Define trainControl with these custom indices using method "cv" with number of folds equal to number of participants
ctrl <- trainControl(method = "cv", index = loo_train_indices,
                       savePredictions = "final")

# Now train your random forest model using these indices:
mod_emoperc_nmf2 <- train(
  max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj,
  method = "rf",
  trControl = ctrl,
  ntree = 500
)



# Access the cross-validation predictions 

# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels <- c("angry", "happy", "other", "sad")  # adjust as needed

mod_emoperc_nmf2_pred_dta <- mod_emoperc_nmf2$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels),
    obs = factor(obs, levels = all_levels)
  )

confusionMatrix(mod_emoperc_nmf2_pred_dta$pred, mod_emoperc_nmf2_pred_dta$obs)
#good


# 0.4554 NIR

mod_emoperc_nmf2$pred%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(Resample), accuracy))+
  geom_point()+
  geom_hline(yintercept = 0.4554)+
  geom_bar(stat = "identity", width = .1)+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()

(42-6)/42

#0.8571429 we show accuracy in 85%^


conf_df <- as.data.frame(confusionMatrix(mod_emoperc_nmf2_pred_dta$pred, mod_emoperc_nmf2_pred_dta$obs)$table)

ggplot(conf_df, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black", size = 4) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
       x = "True Label",
       y = "Predicted Label") +
  theme_minimal()


```

       
Confusion Matrix and Statistics

          Reference
Prediction angry happy other  sad
     angry   271   136   160  124
     happy   148  1412   343  251
     other    37    66    33   53
     sad     131   107   180  327

Overall Statistics
                                               
               Accuracy : 0.5406               
                 95% CI : (0.5246, 0.5566)     
    No Information Rate : 0.4554               
    P-Value [Acc > NIR] : < 0.00000000000000022
                                               
                  Kappa : 0.3073               
                                               
 Mcnemar's Test P-Value : < 0.00000000000000022

Statistics by Class:

                     Class: angry Class: happy Class: other Class: sad
Sensitivity               0.46167       0.8205     0.046089    0.43311
Specificity               0.86842       0.6395     0.949070    0.86177
Pos Pred Value            0.39219       0.6555     0.174603    0.43893
Neg Pred Value            0.89767       0.8098     0.809749    0.85893
Prevalence                0.15533       0.4554     0.189468    0.19979
Detection Rate            0.07171       0.3736     0.008732    0.08653
Detection Prevalence      0.18285       0.5700     0.050013    0.19714
Balanced Accuracy         0.66505       0.7300     0.497579    0.64744

# now we do without other

```{r}

set.seed(44)


# Define trainControl with these custom indices using method "cv" with number of folds equal to number of participants
ctrl <- trainControl(method = "cv", index = loo_train_indices,
                       savePredictions = "final")

# Now train your random forest model using these indices:
mod_emoperc_nmf2_n_oth <- train(
  max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj,
  method = "rf",
  trControl = ctrl,
  ntree = 500
)



# Access the cross-validation predictions 

# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels_noth <- c("angry", "happy", "sad")  # adjust as needed

mod_emoperc_nmf2_n_oth_pred_dta <- mod_emoperc_nmf2_n_oth$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels_noth),
    obs = factor(obs, levels = all_levels_noth)
  )

confusionMatrix(mod_emoperc_nmf2_n_oth_pred_dta$pred, mod_emoperc_nmf2_n_oth_pred_dta$obs)
#good


# 0.5491 NIR

mod_emoperc_nmf2_n_oth_pred_dta%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(Resample), accuracy))+
  geom_point()+
  geom_hline(yintercept = 0.5491)+
  geom_bar(stat = "identity", width = .1)+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()

(42-6)/42

#0.8571429 we show accuracy in 85%^


as.data.frame(confusionMatrix(mod_emoperc_nmf2_n_oth_pred_dta$pred, mod_emoperc_nmf2_n_oth_pred_dta$obs)$table)%>%
ggplot(aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), color = "black", size = 4) +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
       x = "True Label",
       y = "Predicted Label") +
  theme_minimal()


```

# Split by expression only and emotive speech

-   might also look at regular loo, but its bound to be higher in accuracy

# posed LOO - participant

```{r}
dta_scirep_all_wd_by_sbj_pos
unique(dta_scirep_all_wd_by_sbj_pos$posed_spoken)

participants_pos <- unique(dta_scirep_all_wd_by_sbj_pos$participant_private_id)

# Create a list of training indices for leave-one-participant-out:
# For each participant, the training set consists of all rows not belonging to that participant.
loo_train_indices_pos <- lapply(participants_pos, function(p) {
  which(dta_scirep_all_wd_by_sbj_pos$participant_private_id != p)
})
names(loo_train_indices_pos) <- participants_pos

# Define trainControl with these custom indices using method "cv" with number of folds equal to number of participants
ctrl_pos <- trainControl(method = "cv", index = loo_train_indices_pos,
                       savePredictions = "final")

# Now train your random forest model using these indices:
mod_emoperc_nmf2_pos <- train(
  max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj_pos,
  method = "rf",
  trControl = ctrl_pos,
  ntree = 500
)

# Access the cross-validation predictions 

# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels <- c("angry", "happy", "other", "sad")  # adjust as needed

mod_emoperc_nmf2_pos_pred_dta <- mod_emoperc_nmf2_pos$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels),
    obs = factor(obs, levels = all_levels)
  )

confusionMatrix(mod_emoperc_nmf2_pos_pred_dta$pred, mod_emoperc_nmf2_pos_pred_dta$obs)
#good


# 0.4554 NIR

mod_emoperc_nmf2_pos_pred_dta%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(Resample), accuracy))+
  geom_point()+
  # geom_hline(yintercept = 0.4554)+
  
  geom_hline(yintercept = 0.4554, colour = "red", linetype = "dashed")+
  geom_hline(yintercept = .33, colour = "pink", linetype = "dashed")+
  geom_bar(stat = "identity", width = .1)+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()




# as.data.frame(confusionMatrix(mod_emoperc_nmf2_pos_pred_dta$pred, 
#                               mod_emoperc_nmf2_pos_pred_dta$obs)$table)%>%
# 
# ggplot(aes(x = Reference, y = Prediction, fill = Freq)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = Freq), color = "black", size = 4) +
#   scale_fill_gradient(low = "white", high = "steelblue") +
#   labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
#        x = "True Label",
#        y = "Predicted Label") +
#   theme_minimal()




```

pos no other

```{r}


# Now train your random forest model using these indices:
mod_emoperc_nmf2_pos_no_oth <- train(
  max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj_pos,
  method = "rf",
  trControl = ctrl_pos,
  ntree = 500
)

# Access the cross-validation predictions 

# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels_noth 

mod_emoperc_nmf2_pos_no_oth_pred_dta <- mod_emoperc_nmf2_pos_no_oth$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels_noth),
    obs = factor(obs, levels = all_levels_noth)
  )

confusionMatrix(mod_emoperc_nmf2_pos_no_oth_pred_dta$pred, mod_emoperc_nmf2_pos_no_oth_pred_dta$obs)
#good


# 0.5492 NIR
reviews$loo_pos_cv_acc<-

mod_emoperc_nmf2_pos_no_oth_pred_dta%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(as.numeric(factor(Resample))), accuracy))+
  geom_point(size = 1.5)+
    geom_bar(stat = "identity", width = .2)+
  # geom_hline(yintercept = 0.5492)+
   geom_hline(yintercept = 0.5492, colour = "red", linetype = "dashed",size = 1)+
  geom_hline(yintercept = .33, colour = "pink", linetype = "dotted",size = 1)+

  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()+
    ylim(0,.85)+
  ggtitle("expression only")+
  p$graphstyle_int+
    theme(axis.text.x = element_text(size = 6))


reviews$loo_pos_cv_acc

# as.data.frame(confusionMatrix(mod_emoperc_nmf2_pos_no_oth_pred_dta$pred, 
#                               mod_emoperc_nmf2_pos_no_oth_pred_dta$obs)$table)%>%
#   group_by(Reference)%>%
#   mutate(prop = Freq/sum(Freq))%>%
# 
# ggplot(aes(x = Reference, y = Prediction, fill = prop)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = prop), color = "black", size = 4) +
#   scale_fill_gradient(low = "white", high = "steelblue") +
#   labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
#        x = "True Label",
#        y = "Predicted Label") +
#   theme_minimal()



```

loo spk

```{r}
dta_scirep_all_wd_by_sbj_spk


participants_spk <- unique(dta_scirep_all_wd_by_sbj_spk$participant_private_id)

# Create a list of training indices for leave-one-participant-out:
# For each participant, the training set consists of all rows not belonging to that participant.
loo_train_indices_spk <- lapply(participants_spk, function(p) {
  which(dta_scirep_all_wd_by_sbj_spk$participant_private_id != p)
})
names(loo_train_indices_spk) <- participants_spk

# Define trainControl with these custom indices using method "cv" with number of folds equal to number of participants
ctrl_spk <- trainControl(method = "cv", index = loo_train_indices_spk,
                       savePredictions = "final")

# Now train your random forest model using these indices:
mod_emoperc_nmf2_spk <- train(
  max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj_spk,
  method = "rf",
  trControl = ctrl_spk,
  ntree = 500
)

# Access the cross-validation predictions 

# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels 
 
mod_emoperc_nmf2_spk_pred_dta <- mod_emoperc_nmf2_spk$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels),
    obs = factor(obs, levels = all_levels)
  )

confusionMatrix(mod_emoperc_nmf2_spk_pred_dta$pred, mod_emoperc_nmf2_spk_pred_dta$obs)
#  good


# 0.4545 NIR

mod_emoperc_nmf2_spk_pred_dta%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
  ggplot(aes(factor(Resample), accuracy))+
  geom_point()+
  geom_bar(stat = "identity", width = .1)+
  geom_hline(yintercept = 0.4554, colour = "red", linetype = "dashed")+
  geom_hline(yintercept = .33, colour = "pink", linetype = "dashed")+
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  theme_minimal()




# as.data.frame(confusionMatrix(mod_emoperc_nmf2_spk_pred_dta$pred, 
#                               mod_emoperc_nmf2_spk_pred_dta$obs)$table)%>%
#   group_by(Reference)%>%
#   mutate(prop = Freq/sum(Freq))%>%
# ggplot(aes(x = Reference, y = Prediction, fill = prop )) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = prop), color = "black", size = 4) +
#   scale_fill_gradient(low = "white", high = "steelblue") +
#   labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
#        x = "True Label",
#        y = "Predicted Label") +
#   theme_minimal()


```

spk no other

```{r}
# Now train your random forest model using these indices:
mod_emoperc_nmf2_spk_no_oth <- train(
  max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
  data = dta_scirep_all_wd_by_sbj_spk,
  method = "rf",
  trControl = ctrl_spk,
  ntree = 500
)

# Access the cross-validation predictions 
# 2. Compute the overall confusion matrix (ensure that both factors have the same levels)
all_levels_noth 

mod_emoperc_nmf2_spk_no_oth_pred_dta <- mod_emoperc_nmf2_spk_no_oth$pred %>%
  mutate(
    pred = factor(pred, levels = all_levels_noth),
    obs = factor(obs, levels = all_levels_noth)
  )

confusionMatrix(mod_emoperc_nmf2_spk_no_oth_pred_dta$pred, mod_emoperc_nmf2_spk_no_oth_pred_dta$obs)
#good


# 0.5492 NIR

reviews$loo_spk_cv_acc<-
mod_emoperc_nmf2_spk_no_oth_pred_dta%>%
  group_by(Resample) %>%
  summarise(accuracy = mean(pred == obs))%>%
    ggplot(aes(factor(as.numeric(factor(Resample))), accuracy))+
  geom_point(size = 1.5)+
  # geom_hline(yintercept = 0.5492)+
      geom_bar(stat = "identity", width = .2)+
    geom_hline(yintercept = 0.549, colour = "red", linetype = "dashed",size = 1)+

    geom_hline(yintercept = .33, colour = "pink", linetype = "dotted",size = 1)+
  
  labs(title = "Accuracy by Left-Out Participant (Fold)",
       x = "Fold (Left-Out Participant)",
       y = "Accuracy") +
  ggtitle("emotive speech")+
  theme_minimal()+
  ylim(0,.85)+
  p$graphstyle_int+
  theme(axis.title.y = element_blank(),axis.text.x = element_text(size = 6))



reviews$loo_pos_spk_cv_acc <- reviews$loo_pos_cv_acc| reviews$loo_spk_cv_acc

reviews$loo_pos_spk_cv_acc

ggsave("sm_loo_pos_spk_cv_acc.tiff",
       device = 'tiff', reviews$loo_pos_spk_cv_acc,,
       width = 13,
       height = 4,
       dpi = 300,
       compression = "lzw")


# as.data.frame(confusionMatrix(mod_emoperc_nmf2_spk_no_oth_pred_dta$pred, 
#                               mod_emoperc_nmf2_spk_no_oth_pred_dta$obs)$table)%>%
#   group_by(Reference)%>%
#   mutate(prop = Freq/sum(Freq))%>%
# 
# ggplot(aes(x = Reference, y = Prediction, fill = prop)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = prop), color = "black", size = 4) +
#   scale_fill_gradient(low = "white", high = "steelblue") +
#   labs(title = "Confusion Matrix (Leave-One-Participant-Out)",
#        x = "True Label",
#        y = "Predicted Label") +
#   theme_minimal()



```


fetures
```{r}
# dta_rslt_spk_cmfts <- readRDS("~Data/ExportedData/Processeddata/dta_rslt_spk_cmfts.rds")

dta_rslt_spk_cmfts


colnames(dta_rslt_spk_cmfts)

dta_rslt_spk_cmfts[,1:38]
```

to do

just regular loo for completeness

# just take sway too long

```{r eval=FALSE, include=FALSE}
dta_scirep_all_wd_by_sbj

# predict participants ratings from spatiotemp comp data and LOO
?caret::train()

dta_scirep_all_wd_by_subj$posed_spoken_1_2
set.seed(3)

mod_emoperc_nmf_sbj<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj$pred$pred, 
                reference = mod_emoperc_nmf_sbj$pred$obs)


mod_emoperc_nmf_sbj_no_oth<- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)


# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj_no_oth$pred$pred, 
                reference = mod_emoperc_nmf_sbj_no_oth$pred$obs)
#sign

# split bu pos and spk
# pos
mod_emoperc_nmf_sbj_pos<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj_pos,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj_pos$pred$pred, 
                reference = mod_emoperc_nmf_sbj_pos$pred$obs)


mod_emoperc_nmf_sbj_no_oth_pos<- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj_pos,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)


# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj_no_oth_pos$pred$pred, 
                reference = mod_emoperc_nmf_sbj_no_oth_pos$pred$obs)


# spk

mod_emoperc_nmf_sbj_spk<- train(max_expression_resp2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj_spk,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)



# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj_spk$pred$pred, 
                reference = mod_emoperc_nmf_sbj_spk$pred$obs)


mod_emoperc_nmf_sbj_no_oth_spk<- train(max_expression_resp_n_oth2 ~ comp1_scale + comp2_scale + comp3_scale,
                              data = dta_scirep_all_wd_by_sbj_spk,
                              method = "rf",
                              trControl = trainControl(method = "LOOCV"),
                              ntree = 500)


# accuracy metrics
confusionMatrix(data = mod_emoperc_nmf_sbj_no_oth_spk$pred$pred, 
                reference = mod_emoperc_nmf_sbj_no_oth_spk$pred$obs)



```




remove 13017672 from all analysis above

tmp1\<-factor(predict(mod_emoperc_nmf2, dta_rf_test_pos_spk_w_subj%\>% subset(participant_private_id != "13017672")))

tmp2\<-dta_rf_test_pos_spk_w_subj%\>% subset(participant_private_id != "13017672")

tmp2$tmp1<- tmp1
tmp3<- tmp2%>%subset(tmp1!= "other")
confusionMatrix(data = factor(tmp3$tmp1), reference = factor(tmp3\$max_expression_resp2))

```{r}


# Save all specified objects to a file called "project_objects.RData"
save(
  dta_4_plfs,
  dta_4_plfs_sel_long,
  dta_4_plfs_sel_long_test_prct,
  pos_test_files,
  spk_test_files,      # Only include once if there's a duplicate name
  pos_train_files,
  spk_train_files,     # Assuming you meant spk_train_files instead of a repeated spk_test_files
  dta_4_plfs_test_set,
  dta_4_plfs_test_set_agg,
  dta_4_plfs_test_set_stim,
  dta_scirep_grll,
  dta_scirep_grll1,
  dta_rf_test_pos_spk,
  dta_scirep_grll1_wd,
  dta_scirep_grll1_wd_by_vid,
  dta_rf_test_pos_spk_w_subj,
  dta_rf_test_pos_spk_by_vid,
  mod_emoperc_nmf,
  mod_emoperc_nmf2,
  file = "stim_valid_scirep.RData"
)

```

combine environments

```{r}
# Create a new environment to load the saved RData
loaded_env <- new.env()

# Load the RData file into the new environment
load("stim_valid_scirep.RData", envir = loaded_env)

# List the objects in the loaded environment
ls(loaded_env)


# Option 1: Copy all objects from the loaded environment into the current workspace
list2env(as.list(loaded_env), envir = .GlobalEnv)
rm(loaded_env)

# now onwards save sci_rep_response.RData

```
