---
title: "SM"
author: "Helio"
date: "2024-08-23"
output: html_document
---

SM Plots

Fig S1

```{r}
plots_SM$posed_fit_vs_rand <- plot(res_find_k_1,estim_posed_random_2_tbs_5)+
  geom_vline(xintercept = 3)
plots_SM$posed_fit_vs_rand$facet

plots_SM$posed_fit_vs_rand1<- plots_SM$posed_fit_vs_rand$data %>%
    mutate(Data = ifelse(Data == "x", "original", "simulated (permuted)"))%>%
  ggplot(aes(rank, value, colour = Type, shape = Data, linetype = Data ))+
     geom_vline(xintercept = 3,linetype = "dotted")+
  geom_point()+
  geom_line()+
  facet_wrap(~ Measure, scales = "free_y", nrow = 3)+
  # scale_fill_viridis_d(option = "D")
  scale_color_brewer(palette = "Dark2")+# keeps its own range
  xlab("Factorization rank (k) ")+
   ggtitle("NMF component rank validation")+
  
# annotate("text", x = 3, y = Inf, label = "final solution", 
         # vjust = -0.5, angle = 90, 
         # size = 3, colour = "black")+
  theme_light()+
  theme(axis.title.y = element_blank())
   
plots_SM$posed_fit_vs_rand1
facet_grid(pollutant~id, scales = "free_y", labeller = label_both, switch = "y")


SM_Spoken_fit_valid_300dpi.png
ggsave("SM_posed_fit_valid_300dpi.pdf",
       device = 'pdf',    
    plots_SM$posed_fit_vs_rand1,
       width = 7,
       height = 5,
       dpi = 300)

ggsave("SM_posed_fit_valid_300dpi.tiff",
       device = 'tiff',    
     plots_SM$posed_fit_vs_rand1,
        width = 7,
       height = 5,
       dpi = 300)

ggsave("SM_posed_fit_valid_300dpi.png",
       device = 'png',    
      plots_SM$posed_fit_vs_rand1,
              width = 7,
       height = 5,
       dpi = 300)

plots_SM$posed_fit_vs_rand1
```

Fig S2-S3

```{r}



# SM_Spok_consens
NMF::consensusmap(res_find_k_1) #Fig S2

NMF::consensusmap(estim_posed_random_2_tbs_5) #Fig S3


patchwork::wrap_elements(NMF::consensusmap(res_find_k_spk_subs))
 

```

table S 1 and 2

```{r}
write.csv( pos_test_set_res_mts$table, "con_mat_pos.csv")

# round(con_mat_posed$byClass, digits = 2)
# round(con_mat_posed$byClass, digits = 2)
con_mat_pos_bycl_df<-t(as.data.frame(round(pos_test_set_res_mts$byClass,digits = 2)))

pos_test_set_res_mts

con_mat_pos_bycl_df
write.csv(con_mat_pos_bycl_df, "con_mat_pos_bycl_df.csv")

```

Fig S 4

```{r}
plots_SM$spk_fit_vs_rand1<- plots_SM$spk_fit_vs_rand$data %>%
    mutate(Data = ifelse(Data == "x", "original", "simulated (permuted)"))%>%
  ggplot(aes(rank, value, colour = Type, shape = Data, linetype = Data ))+
     geom_vline(xintercept = 3,linetype = "dotted")+
  geom_point()+
  geom_line()+
  facet_wrap(~ Measure, scales = "free_y", nrow = 3)+
  # scale_fill_viridis_d(option = "D")
  scale_color_brewer(palette = "Dark2")+# keeps its own range
  xlab("Factorization rank (k) ")+
   ggtitle("NMF component rank validation")+
  
  theme_light()+
  theme(axis.title.y = element_blank())


plots_SM$spk_fit_vs_rand1
setwd("~/Library/CloudStorage/GoogleDrive-helioclemente.c@gmail.com/My Drive/2022 - University Of Birmingham/HaloStudy/Data/ExportedFigures/SM")

ggsave("SM_Spoken_fit_valid_300dpi.pdf",
       device = 'pdf',    
       plots_SM$spk_fit_vs_rand1,
       width = 7,
       height = 5,
       dpi = 300)

ggsave("SM_Spoken_fit_valid_300dpi.tiff",
       device = 'tiff',    
       plots_SM$spk_fit_vs_rand1,
        width = 7,
       height = 5,
       dpi = 300)

ggsave("SM_Spoken_fit_valid_300dpi.png",
       device = 'png',    
       plots_SM$spk_fit_vs_rand1,
              width = 7,
       height = 5,
       dpi = 300)

plots_SM$spk_fit_vs_rand1
```

fig s5-s6

```{r}
   
# consensus plots save manually at 1200 by 1200 tiff
NMF::consensusmap(res_find_k_spk_subs) #Fig s5
# SM_Spok_consens
NMF::consensusmap(estim_spk_random_1) #Fig S6


 patchwork::wrap_elements(NMF::consensusmap(res_find_k_spk_subs))
 
```

table S3-4

```{r}
ite.csv( spoken_test_set_res_mts_no_neutr$table, "con_mat_spk.csv")

# round(con_mat_posed$byClass, digits = 2)
# round(con_mat_posed$byClass, digits = 2)
con_mat_spk_bycl_df<-t(as.data.frame(round(spoken_test_set_res_mts_no_neutr$byClass,digits = 2)))

spoken_test_set_res_mts_no_neutr

con_mat_spk_bycl_df
write.csv(con_mat_spk_bycl_df, "con_mat_spk_bycl_df.csv")


```

Related to response to reviewers March 2025

move joint NMF here

here we show that the components are similar

```{r}
# single run and single numeric seed for reproducibility

setwd(paths$export_data)

dta_expr_only_4nmf<- readRDS("dta_expr_only_4nmf.rds")

require(tidyverse)
require(NMF)

dta_expr_spk_4nmf <- bind_rows(dta_expr_only_4nmf,
          dta_spk_4nmf)
max(dta_expr_only_4nmf$bin_frame)
max(dta_spk_4nmf$bin_frame)

dta_expr_only_4nmf

colnames(dta_expr_spk_4nmf)

sm<-list()

sm$joint_nmf <- NMF::nmf(dta_expr_spk_4nmf[,6:22], 
                   r = 3, 
                  nrun = 200, 
                  seed=123456,
                 .options = list( 'v')) #


dta_spk_4nmf.rds



# NMF::basismap(sm$joint_nmf) #too long

?NMF::coefmap

NMF::coefmap(sm$joint_nmf, Rowv = F)
NMF::coefmap(mod_rslt_nmfspk,Rowv = F)
NMF::coefmap(mod_rslt_nmfposed, Rowv = F)



sm$joint_nmf
chool_talk$AU_NM
paper_plots$spoken_NMF_hm
paper_plots$







```

```{r}
maxnormalize <- function(x, ...) {
    return((x - min(x, ...)) /(max(x, ...) - min(x, ...))) }




reviews<- list()


reviews$joint_nmf<-
as.data.frame(sm$joint_nmf@fit@H) %>%
  mutate(component = c(1,2,3)) %>%
  gather(AU, coef, -component)%>%
  group_by(as.factor(component))%>%
  mutate(
    # max_comp = max(coef),
         coef = maxnormalize(coef),
         # AU = substring(AU,8,40),
         AU_code = substring(AU,4,5)
         )%>%
  arrange(AU_code)%>%
  ggplot(aes(as.factor(component),AU, fill = coef))+
  geom_tile()+
  theme_classic()+
  # scale_fill_viridis_()
   # scale_fill_()+#)+
   scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  xlab("K")+
  # p$graphstyle_int +
guides(fill=guide_colorbar(ticks.colour = NA))+
  ggtitle("Joint")



# POSED

reviews$pos_nmf<-
as.data.frame(mod_rslt_nmfposed@fit@H) %>%
  mutate(component = c(1,2,3)) %>%
  gather(AU, coef, -component)%>%
  group_by(as.factor(component))%>%
  mutate(
    # max_comp = max(coef),
         coef = maxnormalize(coef),
         # AU = substring(AU,8,40),
         AU_code = substring(AU,4,5)
         )%>%
  arrange(AU_code)%>%
  ggplot(aes(as.factor(component),AU, fill = coef))+
  geom_tile()+
  theme_classic()+
  # scale_fill_viridis_()
   # scale_fill_()+#)+
   scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  xlab("K")+
  # p$graphstyle_int +
guides(fill=guide_colorbar(ticks.colour = NA))+
  ggtitle("Expresion only")

# SPOKEN

reviews$spk_nmf<-
as.data.frame(mod_rslt_nmfspk@fit@H) %>%
  mutate(component = c(2,1,3)) %>%
  gather(AU, coef, -component)%>%
  group_by(as.factor(component))%>%
  mutate(
    # max_comp = max(coef),
         coef = maxnormalize(coef),
         # AU = substring(AU,8,40),
         AU_code = substring(AU,4,5)
         )%>%
  arrange(AU_code)%>%
  ggplot(aes(as.factor(component),AU, fill = coef))+
  geom_tile()+
  theme_classic()+
  # scale_fill_viridis_()
   # scale_fill_()+#)+
   scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  xlab("K")+
  # p$graphstyle_int +
guides(fill=guide_colorbar(ticks.colour = NA))+
    ggtitle("Emotive speech")

#require(patchwork)

reviews$joint_pos_spk_nmf<-
(reviews$joint_nmf+ 
    (reviews$pos_nmf+ theme(axis.title.y = element_blank(),
                            axis.text.y = element_blank()))+
    (reviews$spk_nmf+ theme(axis.title.y = element_blank(),
                            axis.text.y = element_blank())))+
  plot_layout(nrow = 1, ncol = 3, guides = "collect")

reviews$joint_pos_spk_nmf

ggsave(reviews$joint_pos_spk_nmf,
       )

ggsave("sm_joint_and_sprts_nmf.tiff",
       device = 'tiff', reviews$joint_pos_spk_nmf,
       width = 9,
       height = 4,
       dpi = 300,
       compression = "lzw")
  

```

we can see here that the joint NMF capture mostly similar patterns, effectively with three components, where siply component the full one just combines aspects of the otehr two

Now we will plot timeseries from joint split into posed and spoken and compare this to NMF fitted separately on posed and spoken

```{r}
# first store components
dta_expr_spk_4nmf$comp1_k3<- sm$joint_nm@fit@W[,1]
dta_expr_spk_4nmf$comp2_k3<- sm$joint_nm@fit@W[,2]
dta_expr_spk_4nmf$comp3_k3<- sm$joint_nm@fit@W[,3]

unique(dta_expr_spk_4nmf$filename)
# now plot

colnames(dta_expr_spk_4nmf)

# create posed vs spoken idicator variable
dta_expr_spk_4nmf$posed.spoken <- sapply(strsplit(dta_expr_spk_4nmf$filename, "_"), 
                                         function(x) x[2])
unique(dta_expr_spk_4nmf$posed.spoken)

# plot
  dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  # subset(expression!= "neutral")%>%
    # mutate(component = substring(component,4,8))%>%
  
  group_by(posed.spoken, component, bin_frame)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
    group_by(posed.spoken, component)%>%
  
  mutate(
    # coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, coef, color = component))+
  geom_smooth(aes(linetype = component))+
  # geom_smooth(aes(group = subject), se = F)+
  # stat_summary(aes(group = subject, color = component), geom = "smooth")+
  # facet_grid(~expression )+
  theme_classic()+
   xlab("time bin")+
  # p$graphstyle_int+
  scale_color_brewer(palette = "Dark2")+
   # scale_color_viridis_d(option = "magma")+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
    # scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
  theme(legend.position = "top")+
    facet_grid(~posed.spoken)
  
  

# 
  dta_expr_spk_4nmf$posed.spoken <- sapply(strsplit(dta_expr_spk_4nmf$filename, "_"), 
                                         function(x) x[2])
unique(dta_expr_spk_4nmf$posed.spoken)

# plot
  dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  # subset(expression!= "neutral")%>%
    # mutate(component = substring(component,4,8))%>%
  
  group_by(posed.spoken, component, bin_frame, expression)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
    subset(expression!= "neutral")%>%
    group_by(posed.spoken, component)%>%
  
  mutate(
    # coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, coef, color = component))+
  geom_smooth(aes(linetype = component))+
  # geom_smooth(aes(group = subject), se = F)+
  # stat_summary(aes(group = subject, color = component), geom = "smooth")+
  # facet_grid(~expression )+
  theme_classic()+
   xlab("time bin")+
  # p$graphstyle_int+
  scale_color_brewer(palette = "Dark2")+
   # scale_color_viridis_d(option = "magma")+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
    # scale_y_continuous(breaks=c(0,.5,1), limits = c(0, 1))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
  theme(legend.position = "top")+
    facet_grid(posed.spoken~ expression)
  
  
  
#   
#   paper_plots$posed_comp_ts
# paper_plots$posed_comp_ts_hm





```

The joint NMF is overall similar to the separate NMFs

paper_plots\$spoken_ts_across_exp

# heatmap ts

```{r}

  dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  subset(expression!= "neutral")%>%
    mutate(expression = factor(expression, levels = c('angry', 'happy', 'sad')))%>%

  group_by(posed.spoken,component, bin_frame)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
     group_by(posed.spoken,component)%>%
     mutate(coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, component, fill= coef))+
  geom_tile()+
  xlab("time bin")+
      scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  # scale_fill_viridis_c(option = "magma")+
  # geom_smooth(se = F)+
  # facet_grid(~expression)+
  theme_classic()+
  # p$graphstyle_int+
  guides(fill=guide_colorbar(ticks.colour = NA))+
  # theme(panel.spacing = unit(1, "cm"))+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
    facet_grid(~posed.spoken)
  
  
  
# by expression
  
  
  
sm$plt_joint_nmf_ts_hm_by_exp <-
   dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  subset(expression!= "neutral")%>%
    mutate(expression = factor(expression, levels = c('angry', 'happy', 'sad')))%>%

  group_by(posed.spoken,component, bin_frame,expression)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
     group_by(posed.spoken,expression)%>%
     mutate(coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, component, fill= coef))+
  geom_tile()+
  xlab("time bin")+
      scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  # scale_fill_viridis_c(option = "magma")+
  # geom_smooth(se = F)+
  # facet_grid(~expression)+
  theme_classic()+
  # p$graphstyle_int+
  guides(fill=guide_colorbar(ticks.colour = NA))+
  # theme(panel.spacing = unit(1, "cm"))+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
    facet_grid(posed.spoken~expression)



# expression only
sm$plt_joint_nmf_ts_hm_by_exp_pos <-
   dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  subset(expression!= "neutral" & posed.spoken == "posed" )%>%
    mutate(expression = factor(expression, levels = c('angry', 'happy', 'sad')))%>%

  group_by(posed.spoken,component, bin_frame,expression)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
     group_by(posed.spoken,expression)%>%
     mutate(coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  mutate(component = substr(component, 1,5))%>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, component, fill= coef))+
  geom_tile()+
  xlab("time bin")+
      scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  # scale_fill_viridis_c(option = "magma")+
  # geom_smooth(se = F)+
  # facet_grid(~expression)+
  theme_classic()+
  # p$graphstyle_int+
  guides(fill=guide_colorbar(ticks.colour = NA))+
  # theme(panel.spacing = unit(1, "cm"))+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
    facet_grid(~expression)+
  ggtitle("JOINT NMF: expression only")+
  p$graphstyle_int+theme(legend.position = "none")
   
   sm$plt_joint_nmf_ts_hm_by_exp_pos
   
   
   
   # emotive speech
   
sm$plt_joint_nmf_ts_hm_by_exp_spk <-
   dta_expr_spk_4nmf[,c(1:5,23:26)]%>%
  gather(component, coef,-bin_frame, -subject, -expression,-filename,-test.day, -posed.spoken)%>%
  subset(expression!= "neutral" & posed.spoken != "posed" )%>%
    mutate(expression = factor(expression, levels = c('angry', 'happy', 'sad')))%>%

  group_by(posed.spoken,component, bin_frame,expression)%>%

  summarise_if(is.numeric, mean, na.rm = TRUE)%>%
     group_by(posed.spoken,expression)%>%
     mutate(coef_max = max(coef),
         coef = maxnormalize(coef)) %>%
  mutate(component = substr(component, 1,5))%>%
  # ggplot(aes(Dataset, coef , color = Dataset, fill = Dataset))+
  ggplot(aes(bin_frame, component, fill= coef))+
  geom_tile()+
  xlab("time bin")+
      scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  # scale_fill_viridis_c(option = "magma")+
  # geom_smooth(se = F)+
  # facet_grid(~expression)+
  theme_classic()+
  # p$graphstyle_int+
  guides(fill=guide_colorbar(ticks.colour = NA))+
  # theme(panel.spacing = unit(1, "cm"))+
    scale_x_continuous(breaks=c(0,50,100), limits = c(0, 100))+
  theme(panel.spacing.x = unit(1.5, "lines"))+
    facet_grid(~expression)+
  ggtitle("JOINT NMF: emotive speech")+
  p$graphstyle_int+theme(legend.position = "none")

   
   sm$plt_joint_nmf_ts_hm_by_exp_spk
   
   # require(patchwork)
   

   # combine the joint
  sm$plt_joint_nmf_ts_hm_by_exp  <-  (sm$plt_joint_nmf_ts_hm_by_exp_pos/
       sm$plt_joint_nmf_ts_hm_by_exp_spk)
          # plot_layout( guides = "collect")
    # plot_annotation(title = "JOINT NMF")


 sm$ts_hm_pos_spk<-  ((chool_talk$ts_hm+ggtitle("INDIVIDUAL NMFs: expression only")+
                      theme(axis.title.y = element_blank(),
                            axis.text.y = element_blank()))/
   (paper_plots$spoken_NMF_ts_hm+ggtitle("INDIVIDUAL NMFs: emotive speech")+
      theme(axis.title.y = element_blank(),
                            axis.text.y = element_blank())))+
    plot_layout( guides = "collect")
 

 
 # joint export
 
sm$joint_nmf_ind_nmf_hm<-
     (wrap_plots(sm$plt_joint_nmf_ts_hm_by_exp) + 
      wrap_plots(sm$ts_hm_pos_spk))+
     plot_layout(nrow = 1,guides = "collect")+plot_annotation(tag_levels = "A")



ggsave("sm_joint_nmf_ind_nmf_hm.tiff",
       device = 'tiff', sm$joint_nmf_ind_nmf_hm,
       width = 15,
       height = 10,
       dpi = 300,
       compression = "lzw")
```

```{r}

# no data for p9 as he refused to do the task and then day 1 data only for 

# Values to remove
tmp_subj_to_remove <- c(4, 9, 12, 26, 34, 35, 41, 42, 43)




# sd

colnames(dta_expr_spk_4nmf)

reviews$dynm_var<-
dta_expr_spk_4nmf%>%
  select(-comp1_k3, -comp2_k3, -comp3_k3)%>%
  filter(!subject %in% tmp_subj_to_remove)%>%
    subset(expression!= "neutral")%>%
  # group_by(filename, subject,bin_frame, expression, posed.spoken)%>%


  gather("AU", "Coef", -bin_frame,-filename, -expression,-subject,-posed.spoken, -test.day) %>%
      group_by(subject,posed.spoken,AU)%>%
  mutate(Coef = scale(as.numeric(Coef, na.rm = T))) %>%
  

  group_by(bin_frame, expression, posed.spoken,AU)%>%
  subset(!is.na(Coef))%>%
    summarise_at(c("Coef"), sd, na.rm = T) %>%
  
  group_by(expression, posed.spoken,AU)%>%
  # mutate(coef_var = scale(as.numeric(Coef, na.rm = T))) %>%
      mutate(coef_var = normalize_0_1(as.numeric(Coef))) %>%
  mutate(posed.spoken = if_else(posed.spoken == "posed", "expression only","emotive speech"))%>%
  mutate(posed.spoken = factor(posed.spoken, levels = c("expression only","emotive speech")))%>%

  ggplot(aes(bin_frame,AU,fill= coef_var))+
  geom_tile()+
  facet_grid(expression~posed.spoken)+
  # scale_fill_viridis_c(option = "magma")+
    scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  p$graphstyle_int

reviews$dynm_var
  guides(fill=guide_legend(title='variability - sd'))


  
  reviews$dynm_var_mean<-
dta_expr_spk_4nmf%>%
    select(-comp1_k3, -comp2_k3, -comp3_k3)%>%
  filter(!subject %in% tmp_subj_to_remove)%>%
    subset(expression!= "neutral")%>%
  # group_by(filename, subject,bin_frame, expression, posed.spoken)%>%


  gather("AU", "Coef", -bin_frame,-filename, -expression,-subject,-posed.spoken, -test.day) %>%
      group_by(subject,posed.spoken,AU)%>%
  # mutate(sd = scale(as.numeric(Coef, na.rm = T))) %>%
  group_by(bin_frame, expression, posed.spoken,AU)%>%
  subset(!is.na(Coef))%>%
    summarise_at(c("Coef"), mean, na.rm = T) %>%
  group_by(expression, posed.spoken,AU)%>%
  # mutate(coef_mean = scale(as.numeric(Coef, na.rm = T))) %>%
    
      mutate(coef_mean = normalize_0_1(as.numeric(Coef, na.rm = T))) %>%
      mutate(posed.spoken = if_else(posed.spoken == "posed", "expression only","emotive speech"))%>%
  mutate(posed.spoken = factor(posed.spoken, levels = c("expression only","emotive speech")))%>%

  ggplot(aes(bin_frame,AU,fill= coef_mean))+
  geom_tile()+
  facet_grid(expression~posed.spoken)+
      scale_fill_viridis_c(option = "magma", breaks=c(0,.5,1),
                       limits =c(0,1))+#)+
  # scale_fill_viridis_c(option = "magma")+
    p$graphstyle_int

  
  
  reviews$dynm_var_mean <- reviews$dynm_var_mean+theme(axis.text.y = element_blank(),
                                     axis.title.y = element_blank(),
                                     legend.position = "top")
  
   reviews$dynm_var<-  reviews$dynm_var+theme(legend.position = "top")
 
 reviews$dynm_var_mean_patch<-     reviews$dynm_var|reviews$dynm_var_mean 
      
      
      

      ggsave("sm_dynm_var_mean_patch.tiff",
       device = 'tiff',  reviews$dynm_var_mean_patch,
       width = 20,
       height = 15,
       dpi = 300,
       compression = "lzw")

        
         ggsave("sm_dynm_var_mean_patch.tiff",
       device = 'tiff',  reviews$dynm_var_mean_patch,
       width = 30,
       height = 18,
       dpi = 300,
       compression = "lzw")    
         
```
