---
title: "Untitled"
author: "Helio"
date: "2025-07-10"
output: html_document
---



# Load behavioral data

this is data from participants rating the test set for emotion

# rejections

No video face detected in front of the computer

63e6378e04ade071ebf9ff9c

```{r}

library(readr)
dta_scirep_grl <- read_csv("data_exp_220644-v4_task-62d6.csv")

library(readr)
# dta_scirep_grl_h <- read_csv("data-task-62d6_H.csv")
dta_scirep_grll<- dta_scirep_grl%>%
  janitor::clean_names()%>%
  bind_rows(dta_scirep_grl_h%>%janitor::clean_names())


unqiue(dta_scirep_grl_s1$participant_private_id)
dta_scirep_grl_s1<-dta_scirep_grll

# prolific data
library(readr)
dta_scirep_all <- read_csv("data_exp_220644-v6_task-62d6.csv")
View(dta_scirep_all)



dta_scirep_all<- dta_scirep_all%>%
  janitor::clean_names()%>%
  bind_rows(dta_scirep_grll%>%janitor::clean_names())
dta_scirep_grll


unique(dta_scirep_all$participant_private_id)

dta_scirep_all<-dta_scirep_all%>%
  subset(participant_private_id!= "13017685")%>%
  subset(!is.na(participant_private_id))


# select necessary columns and rows (e.g. trial, no practice)

unique(dta_scirep_all$task_version)
unique(dta_scirep_all$participant_private_id)


dta_scirep_all2<- dta_scirep_all%>%
  select(
    #participant_device_type,
         participant_public_id,
         participant_private_id,
          starts_with("spreadsheet"),
         trial_number,
         component_name,
         object_name,
         response)%>%
subset(grepl("Trial",spreadsheet_display))%>%
subset(grepl("Slider",component_name))%>%
  subset(!is.na(response))%>%
  mutate(response = as.numeric(response))

colnames(dta_scirep_all2)

names(dta_scirep_all2)<- str_remove(colnames(dta_scirep_all2), "spreadsheet_")
colnames(dta_scirep_all2)

# quick plots of emotion rating
dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  #subset(object_name!= "other_emotion")%>%
  group_by(expression, posed_spoken)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = object_name,
             shape = object_name))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  geom_point(aes(group = object_name), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1))+
  facet_grid(~posed_spoken)

# quick plots of emotion rating
dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  #mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  #subset(object_name!= "other_emotion")%>%
  group_by(expression, posed_spoken)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = object_name,
             shape = object_name))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  #geom_point(aes(group = object_name), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1))+
  facet_grid(~posed_spoken)





dta_scirep_all2$slider <- str_remove(dta_scirep_all2$object_name, "_slider")

dta_scirep_all2$condition <- if_else(dta_scirep_all2$posed_spoken == "posed", "expression only", "emotive speech")

dta_scirep_all2$condition<- factor(dta_scirep_all2$condition, levels = c("expression only", "emotive speech"))

dta_scirep_all2$filename
reviews$stiom_valid<-

dta_scirep_all2%>%
  #subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  # group_by(participant_private_id,expression, object_name, condition,slider)%>%
  group_by(filename,expression, object_name, condition,slider)%>%  
  
  summarise_if(is.numeric, mean, na.rm = T)%>%
  #mutate(response = if_else(object_name == "other_emotion" & response> 75,NA,response))%>%
  subset(object_name!= "other_emotion")%>%
  group_by(expression, condition)%>%
  mutate(response = scale(response))%>%
  ggplot(aes(expression, response, 
             colour  = slider,
             shape = slider))+
  #geom_boxplot(aes(colour = object_name),  width = .1, position = position_dodge(1))+
  geom_point(aes(group = slider), position = position_dodge(1), alpha = .1)+
    stat_summary(geom = "pointrange",position = position_dodge(1), size = .6)+
  facet_wrap(~condition, scale = "free_y")+
  scale_color_brewer(palette = "Dark2")+
  ylab("response (z)")+
  p$graphstyle_int+
  theme(axis.text.x = element_text(hjust = 1, angle =45),
        legend.position = "top")


reviews$stiom_valid

ggsave("sm_stim_valid.tiff",
       device = 'tiff', reviews$stiom_valid,
       width = 10,
       height = 6,
       dpi = 300,
       compression = "lzw")



```



# plot distributions

outlier tests etc

```{r}


library(dplyr)
library(ggplot2)
library(tidyr)


dta_scirep_all

dta_scirep_all2 %>%
  filter(participant_public_id != "63e6378e04ade071ebf9ff9c") %>%
  group_by(participant_private_id, expression, object_name, posed_spoken) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  mutate(response = if_else(object_name == "other_emotion" & 
                              response > 75, NA_real_, response)) %>%
  #filter(object_name != "other_emotion") %>%
  group_by(expression, posed_spoken) %>%
  mutate(response = scale(response)) %>%
  ggplot(aes(x = expression, y = response, colour = object_name, shape = object_name)) +
    geom_violin(trim=FALSE,position = position_dodge(1))+
  #geom_boxplot(width = 0.1, position = position_dodge(1)) +
  stat_summary(aes(group = object_name), geom = "pointrange", position = position_dodge(1)) +
  facet_wrap(~posed_spoken, nrow = 2) +
  theme_minimal()


colnames(dta_scirep_all2)
unique(dta_scirep_all2$is_outlier)


dta_scirep_all2 <- dta_scirep_all2 %>%
  filter(!is.na(response)) %>%
  group_by(expression, posed_spoken) %>%
  mutate(
    q_1 = quantile(response, 0.25, na.rm = TRUE),
    q_3 = quantile(response, 0.75, na.rm = TRUE),
    iqr =  q_3-q_1,
    lower_bound = q_1 - 1 * iqr,
    upper_bound = q_3 + 1 * iqr,
    is_outlier = response < lower_bound | response > upper_bound
  ) %>%
  ungroup()

unique(dta_scirep_all2$is_outlier)

table(dta_scirep_all2$is_outlier)

dta_scirep_all2$is_outlier2<-NULL
dta_scirep_all2
dta_scirep_all2%>%
  subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  subset(participant_private_id !="13017685")%>%
     # subset(is_outlier == FALSE)%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%

  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
  subset(posed_spoken == "posed")%>%
  
subset(object_name != "other_emotion")%>%
  ggplot(aes(object_name, response, 
             colour  = object_name))+
    stat_summary(geom = "pointrange",
                 position = position_dodge(1))+
  facet_wrap(~expression, scales = "free_y")



dta_scirep_all2%>%
  subset(participant_public_id != "63e6378e04ade071ebf9ff9c")%>%
  subset(participant_private_id !="13017685")%>%
     # subset(is_outlier == FALSE)%>%
  group_by(participant_private_id,expression, object_name, posed_spoken)%>%

  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
  subset(posed_spoken != "posed")%>%
  
subset(object_name != "other_emotion")%>%
  ggplot(aes(object_name, response, 
             colour  = object_name))+
    stat_summary(geom = "pointrange",
                 position = position_dodge(1))+
  facet_wrap(~expression, scales = "free_y")



```

load the test rf data

```{r}
#  dta_pos_cfmts_4_pca_test <- readRDS("~//Processeddata/dta_pos_cfmts_4_pca_test.rds")
# > dta_spoken_cfmts_4_pca_test <- readRDS("~Processeddata/dta_spoken_cfmts_4_pca_test.rds")

unique(dta_pos_cfmts_4_pca_test$filename)
unique(dta_spoken_cfmts_4_pca_test$filename)

dta_pos_cfmts_4_pca_test
dta_spoken_cfmts_4_pca_test

# bind expression only and emotive speech
dta_rf_test_pos_spk<- bind_rows(dta_pos_cfmts_4_pca_test,
          dta_spoken_cfmts_4_pca_test)

# remove patterns to match the filename pattern in the behavioural data
dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,
                                          "./")

dta_rf_test_pos_spk$filename<- str_remove(dta_rf_test_pos_spk$filename,".csv")

# check that we have the same filenames in rf and behavioural data
setdiff(dta_rf_test_pos_spk$filename,dta_scirep_all2$filename)
intersect(dta_rf_test_pos_spk$filename,dta_scirep_all2$filename)



```

# now create categorical ratings for emotion

We will assign a category based on the maximum rating, we can ignore neutral in here since we are focusing on test set, and we already use other to make the rating difficult

```{r}

colnames(dta_scirep_all2)

dta_scirep_all2<- dta_scirep_all2[,c(1:16)]

# Specify the identifier columns – these are all columns you do not wish to pivot.
id_vars <- c("participant_device_type", "participant_public_id", "participant_private_id", 
             "spreadsheet", "display", "filename", 
             "posed_spoken", "day", "subject", "expression",
             "frame", "filename_mp4", "posed_spoken_1_2", 
             "component_name", "object_name")


# Pivot the data: each unique spreadsheet_expression becomes a new column

dta_scirep_all_wd <- dta_scirep_all2 %>%
  pivot_wider(
    #id_cols = all_of(id_vars),
    names_from = object_name,
    values_from = response
  )

dta_scirep_all_wd

# Create a new column 'max_expression' containing the expression with the highest response

selected_cols<- c("happy_slider", "angry_slider", "sad_slider", "other_emotion")

dta_scirep_all_wd <- dta_scirep_all_wd %>%
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()%>%
    rowwise() %>%
  mutate(max_expression_resp_n_oth = {
    values <- c_across(all_of(selected_cols[1:3]))
    selected_cols[1:3][which.max(values)]
  }) %>%
    ungroup()

unique(dta_scirep_all_wd$max_expression_resp_n_oth)
table(dta_scirep_all_wd$max_expression_resp)


#aggregated across subjects


dta_scirep_all_wd_by_vid<-  dta_scirep_all_wd%>%
  group_by(display,
  filename, posed_spoken,expression,day,subject,
  filename_mp4, posed_spoken_1_2)%>%
  summarise_if(is.numeric, mean, na.rm = T)%>%
  ungroup()%>%
#create emotion catgeory by video
  rowwise() %>%
  mutate(max_expression_resp = {
    values <- c_across(all_of(selected_cols))
    selected_cols[which.max(values)]
  }) %>%
  ungroup()%>%
  rowwise() %>%
  mutate(max_expression_resp_n_oth = {
    values <- c_across(all_of(selected_cols[1:3]))
    selected_cols[1:3][which.max(values)]
  }) %>%
    ungroup()

dta_scirep_all_wd_by_vid$max_expression_resp
dta_scirep_all_wd_by_vid$max_expression_resp_n_oth

# now merge the nnf classification data with subject level perceptual data on the test set
dta_scirep_all_wd
dta_scirep_all_wd_by_sbj<- left_join(dta_scirep_all_wd, dta_rf_test_pos_spk,
                                        by = c("filename", "expression"))



# now merge the nmf classification data with subject aggregated  perceptual data on the test set

dta_scirep_all_wd_by_vid


dta_scirep_all_wd_by_vid<- left_join(dta_scirep_all_wd_by_vid, dta_rf_test_pos_spk, 
                                       by = c("filename", "expression"))

# keep just emotion name
dta_scirep_all_wd_by_vid$max_expression_resp2<- sub("_.*", "",  dta_scirep_all_wd_by_vid$max_expression_resp)


dta_scirep_all_wd_by_sbj$max_expression_resp2 <- sub("_.*", "",  dta_scirep_all_wd_by_sbj$max_expression_resp)


dta_scirep_all_wd_by_vid$max_expression_resp_n_oth2<- sub("_.*", "",  dta_scirep_all_wd_by_vid$max_expression_resp_n_oth)


dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth2 <- sub("_.*", "",  dta_scirep_all_wd_by_sbj$max_expression_resp_n_oth)






dta_scirep_all_wd

write_csv(dta_scirep_all_wd, "dta_scirep_all_wd.csv")



dta_scirep_all_wd$max_expression_resp_n_oth




```

Now check correspondence of defined emotion categories and perceptual rating derived categories

```{r}
library(caret)
# check correspondence

options(scipen = 999)

# check factors first
dta_scirep_all_wd_by_vid$prediction<- factor(dta_scirep_all_wd_by_vid$prediction)

dta_scirep_all_wd_by_vid$max_expression_resp2<- factor(dta_scirep_all_wd_by_vid$max_expression_resp2)


dta_scirep_all_wd_by_sbj$max_expression_resp2<- factor(dta_scirep_all_wd_by_sbj$max_expression_resp2)


dta_scirep_all_wd_by_sbj$prediction<- factor(dta_scirep_all_wd_by_sbj$prediction)



```

